---
title: "Hypothesis 3"
description: "We hypothesize that (a) US-based researchers are most likely to use population terms and (b) that US-specific racial and ethnic categories have grown in use in biomedical and diversity abstracts over the past three decadess."
output: html_document
weight: 3
---

To test Hypothesis 3, we evaluated the geographical basis of population terminology use and variation in US-specific racial and ethnic categories over time. First, we extracted and parsed all of the address information of authors in our diversity corpus, recoding these data into ISO Alpha-2 country codes. We then examined how often all population terms (i.e. the dictionary of 8,200+ terms) were used from researchers in each country across both the biomedical and diversity samples. 

```{r prereqs_4ab, message = FALSE, results = FALSE, warning = FALSE, echo=FALSE}



```

```{r figure 3a, echo=FALSE}
figure_3A <- pop_terms_bycountry %>%   
  drop_na(country) %>%
  group_by(year) %>% 
  count(term, country, sort = TRUE) %>% 
  ungroup() %>% 
  filter(term == "all population terms" ) %>% 
  filter(  country == "United States" | country == "China" | country == "United Kingdom" | 
           country == "Germany" | country == "Australia" 
  #         | country == "France" | country == "Canada" | country == "Netherlands" 
  #         | country == "Australia" | country == "Spain"   
           ) %>%  
  ggplot() + geom_line(aes(y = n, x = year, colour = country), size = 1.3, stat="identity") +
  labs(title = "    Figure 3A. Total Growth of All Population Terms \n by Country in the Biomedical Sample (1990-2017)") + 
  theme_minimal() +
  ylab("Term Count") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        #legend.position = "none",
        legend.box = "vertical", 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(values=c("#61B329", "#E2B306", "#3B9AB2", "#E86F00", "#E21F00")) +
  scale_x_continuous(limits = c(1990, 2017),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015)); ggplotly()
```

Figure 3A shows that US-based researchers are using population terms more than all other countries with those based in China, the UK, Australia, and Germany as the next closest countries. While the US totals steadily increase throughout the decade, it seems that all countries, especially those based in the US, have shared a sharp increase in the usage of population terms since until 2013. This spike, however, largely seems to disappear after controlling for the total number of abstracts available after this year. 

```{r figure 3b, echo=FALSE}
# total articles each year 
bycountry_prc_counts <- text_data %>% 
  drop_na(country) %>% 
  group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n) 

bycountry_prc_counts <- pop_terms_bycountry %>% 
  filter(all_pop_terms == 1) %>% 
  group_by(country, year) %>% 
  summarise(cnt_pop_terms = n_distinct(pubmed_id)) %>% 
  right_join(bycountry_prc_counts, by = "year") %>% 
  mutate(prc_pop_terms = round(cnt_pop_terms / total * 100, digits = 2))
# graphing 
figure_3B <- bycountry_prc_counts %>% 
  drop_na(country) %>% 
  filter(  country == "United States" | country == "China" | country == "United Kingdom" | 
           country == "Germany" | country == "Australia" 
  #         | country == "France" | country == "Canada" | country == "Netherlands" 
  #         | country == "Australia" | country == "Spain"   
           ) %>% 
  ggplot() + 
  geom_line(aes(y = prc_pop_terms, x = year, colour = country), size = 1.3, stat="identity") +
  labs(title = "      Figure 3B. Proportional Growth of All Population Terms \n by Country in the Biomedical Sample (1990-2017)",
       color =  "Country") + 
  theme_minimal() +
  ylab("Proportion of Articles") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        #legend.position = "none",
        legend.box = "vertical", 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(values=c("#61B329", "#E2B306", "#3B9AB2", "#E86F00", "#E21F00")) +
  scale_x_continuous(limits = c(1990, 2017),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015)); ggplotly()
```

Figure 3B shows that US-based researchers use population terms in around 10-12% of all abstracts. On the other hand, no other country uses population terms in more than 5% of abstracts in a given year. While the use of population terms seems to be growing slighltly in most other countries, the proportion is declining slighltly in US-based work.

Now that we know US-based researchers are most likely to use population terms (a result we fully expected), we want to further examine how OMB/US Census terms are used in our two corpora. To ensure that we are capturing US-centric terminology, we [compiled a dictionary](https://growthofdiversity.netlify.app/analyses/dictionaries) used to catch to the main OMB/US Census terms, including Asian, White, Black, Native American, Pacific Islander, Hispanic/Latinx as well as more general terms like race and ethnicity. Given that these terms can also refer to populations outside of the US, we created bigrams of US-centric terms to ensure they were specific to the American context (i.e. we constructed bigrams of all adjacent uses of African and American). 

```{r prereqs_3c, message = FALSE, results = FALSE, warning = FALSE, echo=FALSE}
setwd("~/Documents/Diversity/Data")
bm_text_data <- read_csv("biomedical_corpus.csv") %>% 
  rowid_to_column(var = "id") %>% 
  rename(authors = AU, title = TI, publication = SO, #author_keywords = DE,
         abstract = AB, references = CR, year = PY, times_cited = TC, 
         department = C1, pubmed_id = UT) %>% 
  drop_na(abstract) %>%
  select(id, authors, title, year, publication, abstract, 
         references, year, times_cited, pubmed_id, department) %>% 
# filtering 2018 articles because they seem to be incomplete 
  filter(year != "2018")



# recoding frequently occuring population-related bigrams 
us_pop_bigrams <- bm_text_data %>% 
  unnest_tokens(bigram, abstract, token = "ngrams", n = 2) %>% 
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = black_bigrams), yes = "african-american", no = bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = native_american_bigrams), yes = "native-american", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = pacific_islander_bigrams), yes = "pacific-islander", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = hispanic_american_bigrams), yes = "hispanic-american", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = asian_american_bigrams), yes = "asian-american", no = rc_bigram)) %>%
  select(-bigram) %>% 
  filter(rc_bigram == "african-american" | 
         rc_bigram == "native-american" | 
         rc_bigram == "hispanic-american" |
         rc_bigram == "asian-american" |
         rc_bigram == "pacific-islander") %>% 
  count(rc_bigram, year, sort = TRUE) %>% 
  rename(word = rc_bigram)

# tokenizing the abstract data into words 
bm_abstract_data <- bm_text_data %>% 
  unnest_tokens(word, abstract) #%>% anti_join(stop_words)
# adding custom set of stopwords 
my_stopwords <- tibble(word = c(as.character(1:14), "[0-9]", 
                                "rights", "reserved", "copyright", "elsevier", 
                                "women.recommendations", 
                                "cs.man.ac.uk", "sexes.please", "movements.baby"))
bm_abstract_data <- bm_abstract_data %>% 
  anti_join(stop_words) %>% 
  anti_join(my_stopwords); rm(my_stopwords)

# combining the hyphenated words to our full dataset 
us_pop_counts <- bm_abstract_data %>% 
  group_by(year) %>% 
  count(word, sort = TRUE) %>% 
  ungroup() %>% 
  select(word, year, n) %>% 
  bind_rows(us_pop_bigrams) %>% 
  arrange(-n)
```

```{r figure 3c, message = FALSE, warning = FALSE, echo=FALSE}
setwd("~/Documents/Diversity/Data")
h1_dictionary <- read_csv("diversity_project - h1_dictionary.csv")
h3_dictionary <- read_csv("diversity_project - h3_dictionary.csv")
omb_black <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$black), "z3x)\\b"), collapse = "|")
omb_native_american <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$native_american), "z3x)\\b"), collapse = "|")
omb_pacific_islander <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$pacific_islander), "z3x)\\b"), collapse = "|")
omb_hispanic_latinx <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$hispanic_latinx), "z3x)\\b"), collapse = "|")
omb_asian <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$asian), "z3x)\\b"), collapse = "|")
omb_white <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$white), "z3x)\\b"), collapse = "|")
omb_racial <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$race), "z3x)\\b"), collapse = "|")
omb_ethnicity <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$ethnicity), "z3x)\\b"), collapse = "|")
omb_diversity <- paste(c("\\b(?i)(z3x",na.omit(h1_dictionary$diversity), "z3x)\\b"), collapse = "|")

# graphing us-specific terms over time using census categories 
# https://www.census.gov/topics/population/race/about.html
figure_3C <- us_pop_counts %>% 
  mutate(term = ifelse(test = str_detect(string = word,
                       pattern = omb_black), yes = "black", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_white), yes = "white", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_asian), yes = "asian", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                      pattern = omb_hispanic_latinx), yes = "hispanic/latinx", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_racial), yes = "race", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_ethnicity), yes = "ethnicity", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_native_american), yes = "native-american", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_pacific_islander), yes = "pacific-islander", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_diversity), yes = "diversity", no = term)) %>%
  filter(term == "black" | term == "white" | term == "asian" |
         term == "hispanic/latinx" | term == "race" | term == "ethnicity" |
         term == "native-american" | term == "pacific-islander" | term == "diversity" ) %>%  
  group_by(term, year) %>% summarise(n = sum(n)) %>% arrange(-n) %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), size = 1.3, stat="identity") + 
  theme_minimal() +
  labs(title = "     Figure 3C. Total Growth of OMB/US Census Terms \n in Biomedical Sample (1990-2017)",
       color =  "Term") + 
  ylab("Term Count") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(labels=c("asian","black","diversity","ethnicity","hispanic/latinx",
                              "native-american","pacific-islander","race","white"),
                     values=c("#E86F00", "#3B9AB2", "#000000", "#E21F00", "#00468b", 
                              "#6b0f6b", "#d23ccc",  "#61B329", "#E2B306")) +
  scale_x_continuous(limits = c(1990, 2017),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015)); ggplotly() 
```

Figure 3C shows the OMB US Census terms are growing over time, but this growth is quite subdued compared to our previous results. For example, no term set exceeds more than 400 uses in a given year. Of these category sets, ethnicity grows the most followed by Asian, White, race, and Black. 

```{r prereqs_figure_3D, message = FALSE, results = FALSE, warning = FALSE, echo=FALSE}

setwd("~/Documents/Diversity/Data")
# loading the .csv file 
text_data_3d <- read_csv("biomedical_corpus.csv") %>% 
  rowid_to_column(var = "id") %>% 
  rename(authors = AU, title = TI, publication = SO, #author_keywords = DE,
         abstract = AB, references = CR, year = PY, times_cited = TC, 
         department = C1, pubmed_id = UT) %>% 
  #drop_na(abstract) %>%
  select(id, authors, title, year, publication, abstract, 
         references, year, times_cited, pubmed_id, department) %>% 
# filtering 2018 articles because they seem to be incomplete 
  filter(year != "2018")

abstract_data_3d <- text_data_3d %>% 
  unnest_tokens(word, abstract) %>% 
  anti_join(stop_words)

# adding custom set of stopwords 
my_stopwords <- tibble(word = c(as.character(1:14), "[0-9]", 
                                "rights", "reserved", "copyright", "elsevier"))
abstract_data_3d <- abstract_data_3d %>% 
  anti_join(my_stopwords); rm(my_stopwords)

# graphing us-specific terms over time 
us_pop_counts_3d <- abstract_data_3d %>% 
  mutate(term = ifelse(test = str_detect(string = word,
                       pattern = omb_black), yes = "black", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_white), yes = "white", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_asian), yes = "asian", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                      pattern = omb_hispanic_latinx), yes = "hispanic/latinx", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_racial), yes = "race", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_ethnicity), yes = "ethnicity", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_native_american), yes = "native-american", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_pacific_islander), yes = "pacific-islander", no = term)) %>% 
  mutate(black = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(black)\\b"), yes = 1, no = 0)) %>% 
  mutate(white = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(white)\\b"), yes = 1, no = 0)) %>%
  mutate(asian = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(asian)\\b"), yes = 1, no = 0)) %>%
  mutate(hispanic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(hispanic/latinx)\\b"), yes = 1, no = 0)) %>% 
  mutate(native_american = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(native-american)\\b"), yes = 1, no = 0)) %>% 
  mutate(pacific_islander = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(pacific-islander)\\b"), yes = 1, no = 0)) %>% 
  mutate(racial = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(race)\\b"), yes = 1, no = 0)) %>% 
  mutate(ethnic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(ethnicity)\\b"), yes = 1, no = 0)) %>% 
  mutate(diversity = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(diversity)\\b"), yes = 1, no = 0))

# total articles each year 
us_pop_prc_counts <- text_data_3d %>% 
  #drop_na(abstract) %>% 
  group_by(year) %>% 
  count(year) %>% 
  ungroup() %>% 
  rename(total = n)  

# articles with term mentioned each year 
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(black == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_black = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_black = round(cnt_black / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(white == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_white = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_white = round(cnt_white / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(asian == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_asian = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_asian = round(cnt_asian / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(hispanic == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_hispanic = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_hispanic = round(cnt_hispanic / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(native_american == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_native_american = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_native_american = round(cnt_native_american / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(pacific_islander == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_pacific_islander = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_pacific_islander = round(cnt_pacific_islander / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(racial == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_racial = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_racial = round(cnt_racial / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(ethnic == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_ethnic = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_ethnic = round(cnt_ethnic / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts_3d %>% 
  filter(diversity == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_diversity = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_diversity = round(cnt_diversity / total * 100, digits = 2))

### THIS IS BROKEN HERE AND I HAVE NO IDEA WHYYYYYY


```

```{r figure_3D, message = FALSE, warning=FALSE, echo=FALSE}
# graphing terms over time 
figure_3D <- us_pop_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_black, x = year, colour = "black"), size = 1.3, stat="identity") +
  geom_line(aes(y = prc_white, x = year, colour = "white"),size = 1.3, stat="identity") +
  geom_line(aes(y = prc_asian, x = year, colour = "asian"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_hispanic, x = year, colour = "hispanic/latinx"),size = 1.3, stat="identity") +
  geom_line(aes(y = prc_native_american, x = year, colour = "native-american"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_pacific_islander, x = year, colour = "pacific-islander"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_ethnic, x = year, colour = "ethnicity"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_racial, x = year, colour = "race"), size = 1.3, stat="identity") +
  geom_line(aes(y = prc_diversity, x = year, colour = "diversity"), size = 1.3, stat="identity") +
  labs(title = "    Figure 3D. Proportional Growth of \n US-Specific Terms in Biomedical Sample (1990-2017)",
       color =  "Term") + 
  theme_minimal() +
  ylab("Proportion of Articles") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        #legend.position = "none",
        legend.box = "vertical", 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(labels=c("asian","black","diversity","ethnicity","hispanic/latinx",
                              "native-american","pacific-islander","race","white"),
                     values=c("#E86F00", "#3B9AB2", "#000000", "#E21F00", "#00468b", 
                              "#6b0f6b", "#d23ccc",  "#61B329", "#E2B306")) +
  scale_x_continuous(limits = c(1990, 2017),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015)); ggplotly()
```

Similarly, Figure 3D shows that the OMB/US Census terms vary little over time. While there are some brief spikes in usage, most totals never surpass 1% of abstracts in any given year. This plot further corroborates for the notion that researchers are moving away from the use of racial and ethnic terms to describe populations in their work.

```{r prereqs_3e, message = FALSE, results = FALSE, warning = FALSE, echo=FALSE}
for (pkg in c("tidyverse", "igraph", "tidytext", "ggplot2", "DT",
              "grid", "gridExtra", "reshape2", "extrafont",
               "plotly", "stringi", "networkD3")) {library(pkg, character.only = TRUE)}

setwd("~/Documents/Diversity/Data")
# loading the .csv file 
text_data_div <- read_csv("diversity_corpus.csv") %>% 
# filtering 2018 articles because they seem to be incomplete 
  filter(year != "2018") 

black_bigrams <- "\\b(?i)(african american|african americans|afro american|afro americans|african-american|african-americans|afro-american|afro-americans)\\b"
native_american_bigrams <- "\\b(?i)(native american|native americans|american indian|american indians|alaskan native|alaskan natives|alaska nativez|alaska natives|native-american|native-americans|american-indian|american-indians|alaskan-native|alaskan-natives|alaska-native|alaska-natives)\\b"
pacific_islander_bigrams <- "\\b(?i)(pacific islander|pacific islanders|native hawaiian|native hawaiians|pacific-islander|pacific-islanders|native-hawaiian|native-hawaiians)\\b"
hispanic_american_bigrams <- "\\b(?i)(hispanic-american|hispanic american|hispanic-americans|hispanic americans)\\b"
asian_american_bigrams <- "\\b(?i)(asian-american|asian american)\\b"
asian_american_bigrams <- "\\b(?i)(caucasian-american|caucasian american)\\b"

# recoding frequently occuring population-related bigrams 
us_pop_bigrams_div <- text_data_div %>% 
  unnest_tokens(bigram, abstract, token = "ngrams", n = 2) %>% 
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = black_bigrams), yes = "african-american", no = bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = native_american_bigrams), yes = "native-american", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = pacific_islander_bigrams), yes = "pacific-islander", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = hispanic_american_bigrams), yes = "hispanic-american", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
         pattern = asian_american_bigrams), yes = "asian-american", no = rc_bigram)) %>%
  select(-bigram) %>% 
  filter(rc_bigram == "african-american" | 
         rc_bigram == "native-american" | 
         rc_bigram == "hispanic-american" |
         rc_bigram == "asian-american" |
         rc_bigram == "pacific-islander") %>% 
  count(rc_bigram, year, sort = TRUE) %>% 
  rename(word = rc_bigram)

# tokenizing the abstract data into words 
abstract_data_div <- text_data_div %>% 
  unnest_tokens(word, abstract) #%>% anti_join(stop_words)
# adding custom set of stopwords 
my_stopwords <- tibble(word = c(as.character(1:14), "[0-9]", 
                                "rights", "reserved", "copyright", "elsevier", 
                                "women.recommendations", 
                                "cs.man.ac.uk", "sexes.please", "movements.baby"))
abstract_data_div <- abstract_data_div %>% 
  anti_join(stop_words) %>% 
  anti_join(my_stopwords); rm(my_stopwords)

# combining the hyphenated words to our full dataset 
us_pop_counts_div <- abstract_data_div %>% 
  group_by(year) %>% 
  count(word, sort = TRUE) %>% 
  ungroup() %>% 
  select(word, year, n) %>% 
  bind_rows(us_pop_bigrams_div) %>% 
  arrange(-n)
```

```{r figure 3e, echo=FALSE}
setwd("~/Documents/Diversity/Data")
h3_dictionary <- read_csv("diversity_project - h3_dictionary.csv") 
omb_black <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$black), "z3x)\\b"), collapse = "|")
omb_native_american <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$native_american), "z3x)\\b"), collapse = "|")
omb_pacific_islander <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$pacific_islander), "z3x)\\b"), collapse = "|")
omb_hispanic_latinx <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$hispanic_latinx), "z3x)\\b"), collapse = "|")
omb_asian <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$asian), "z3x)\\b"), collapse = "|")
omb_white <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$white), "z3x)\\b"), collapse = "|")
omb_racial <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$race), "z3x)\\b"), collapse = "|")
omb_ethnicity <- paste(c("\\b(?i)(z3x",na.omit(h3_dictionary$ethnicity), "z3x)\\b"), collapse = "|")
omb_diversity <- paste(c("\\b(?i)(z3x",na.omit(h1_dictionary$diversity), 
                         "z3x)\\b"), collapse = "|")

# graphing us-specific terms over time using census categories 
# https://www.census.gov/topics/population/race/about.html
figure_3E <- us_pop_counts_div %>% 
  mutate(term = ifelse(test = str_detect(string = word,
                       pattern = omb_black), yes = "black", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_white), yes = "white", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_asian), yes = "asian", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                      pattern = omb_hispanic_latinx), yes = "hispanic/latinx", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_racial), yes = "race", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_ethnicity), yes = "ethnicity", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_native_american), yes = "native-american", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_pacific_islander), yes = "pacific-islander", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_diversity), yes = "diversity", no = term)) %>%
  filter(term == "black" | term == "white" | term == "asian" |
         term == "hispanic/latinx" | term == "race" | term == "ethnicity" |
         term == "native-american" | term == "pacific-islander" | term == "diversity") %>%  
  group_by(term, year) %>% summarise(n = sum(n)) %>% arrange(-n) %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), size = 1.3, stat="identity") + 
  theme_minimal() +
  labs(title = "     Figure 3E. Total Growth of OMB/US Census \n Terms in Diversity Sample (1990-2017)",
       color =  "Term") + 
  ylab("Term Count") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(labels=c("asian","black","diversity","ethnicity","hispanic/latinx",
                              "native-american","pacific-islander","race","white"),
                     values=c("#E86F00", "#3B9AB2", "#000000", "#E21F00", "#00468b", 
                              "#6b0f6b", "#d23ccc",  "#61B329", "#E2B306")) +
  scale_x_continuous(limits = c(1990, 2017),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015)); ggplotly() 
```

Like we saw in Hypothesis 2, Figure 3E shows that the trends for all ategories in the diversity sample are inflated compared to the biomedical sample. In addition to having counts nearly twice the biomedical sample, ethnicity and race are the top two category sets used in this sample with Black, White, Asian and Hispanic/Latinx all trailing closely behind.  

```{r prereqs_figure_3F, message = FALSE, results = FALSE, echo=FALSE}
# combining the hyphenated words to our full dataset 
us_pop_counts_3f <- abstract_data_div %>% 
  group_by(year) %>% 
  count(word, pubmed_id, sort = TRUE) %>% 
  ungroup() %>% 
  select(word, pubmed_id, year, n) %>% 
  bind_rows(us_pop_bigrams_div) 

# graphing us-specific terms over time 
us_pop_counts_3f <- us_pop_counts_3f %>% 
  mutate(term = ifelse(test = str_detect(string = word,
                       pattern = omb_black), yes = "black", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_white), yes = "white", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_asian), yes = "asian", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                      pattern = omb_hispanic_latinx), yes = "hispanic/latinx", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_racial), yes = "race", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_ethnicity), yes = "ethnicity", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_native_american), yes = "native-american", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = omb_pacific_islander), yes = "pacific-islander", no = term)) %>% 
  mutate(black = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(black)\\b"), yes = 1, no = 0)) %>% 
  mutate(white = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(white)\\b"), yes = 1, no = 0)) %>%
  mutate(asian = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(asian)\\b"), yes = 1, no = 0)) %>%
  mutate(hispanic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(hispanic/latinx)\\b"), yes = 1, no = 0)) %>% 
  mutate(native_american = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(native-american)\\b"), yes = 1, no = 0)) %>% 
  mutate(pacific_islander = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(pacific-islander)\\b"), yes = 1, no = 0)) %>% 
  mutate(racial = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(race)\\b"), yes = 1, no = 0)) %>% 
  mutate(ethnic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(ethnicity)\\b"), yes = 1, no = 0)) %>% 
  mutate(ethnic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(diversity)\\b"), yes = 1, no = 0))

# total articles each year 
us_pop_prc_counts_div <- text_data_div %>% 
  group_by(year) %>% 
  count(year) %>% 
  ungroup() %>% 
  rename(total = n)  

# articles with term mentioned each year 
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(black == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_black = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_black = round(cnt_black / total * 100, digits = 2))
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(white == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_white = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_white = round(cnt_white / total * 100, digits = 2))
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(asian == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_asian = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_asian = round(cnt_asian / total * 100, digits = 2))
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(hispanic == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_hispanic = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_hispanic = round(cnt_hispanic / total * 100, digits = 2))
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(native_american == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_native_american = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_native_american = round(cnt_native_american / total * 100, digits = 2))
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(pacific_islander == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_pacific_islander = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_pacific_islander = round(cnt_pacific_islander / total * 100, digits = 2))
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(racial == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_racial = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_racial = round(cnt_racial / total * 100, digits = 2))
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(ethnic == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_ethnic = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_ethnic = round(cnt_ethnic / total * 100, digits = 2))
us_pop_prc_counts_div <- us_pop_counts_3f %>% 
  filter(diversity == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_diversity = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts_div, by = "year") %>% 
  mutate(prc_diversity = round(cnt_diversity / total * 100, digits = 2))
```

```{r figure_3F, echo=FALSE}
# graphing terms over time 
figure_3F <- us_pop_prc_counts_div %>% ggplot() + 
  geom_line(aes(y = prc_black, x = year, colour = "black"), 
            size = 1.3, stat="identity") +
  geom_line(aes(y = prc_white, x = year, colour = "white"),
            size = 1.3, stat="identity") +
  geom_line(aes(y = prc_asian, x = year, colour = "asian"), 
            size = 1.3,stat="identity") +
  geom_line(aes(y = prc_hispanic, x = year, colour = "hispanic/latinx"),
            size = 1.3, stat="identity") +
  geom_line(aes(y = prc_native_american, x = year, colour = "native-american"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_pacific_islander, x = year, colour = "pacific-islander"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_ethnic, x = year, colour = "ethnicity"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_racial, x = year, colour = "race"), size = 1.3, stat="identity") +
  geom_line(aes(y = prc_diversity, x = year, colour = "diversity"), size = 1.3, stat="identity") +
  labs(title = "    Figure 3F. Proportional Growth of OMB/US Census \n Terms in Diversity Sample (1990-2017)",
       color =  "Term") + 
  theme_minimal() +
  ylab("Proportion of Articles") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        #legend.position = "none",
        legend.box = "vertical", 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(labels=c("asian","black","diversity","ethnicity","hispanic/latinx",
                              "native-american","pacific-islander","race","white"),
                     values=c("#E86F00", "#3B9AB2", "#000000", "#E21F00", "#00468b", 
                              "#6b0f6b", "#d23ccc",  "#61B329", "#E2B306")) +
  scale_x_continuous(limits = c(1990, 2017),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015)); ggplotly()
```

Finally, Figure 3F exhibits a non-linear trend analogous to the one identified in Hypothesis 1. Most prominently, the term ethnicity rises sharply from 1993 to 2006 before dropping back to about 5% of all published abstracts in the sample. Most of the other terms in this plot (i.e. Asian, White, Hispanic/Latinx, and race) follow the same parabolic curve while others (like Black, Pacific-Islander and Native-American) show almost no variance at all over the course of the sampling period. 

### Main Takeaway 

Overall, Figures 3A-3F support our hypothesis that *(a)* US-based researchers tend to use population terminology at disproportionately higher rates than scientists from other countries and *(b)* that terminology based on the US-Census and OMB Directive 15 have increased since 1990. However, when looking more closely at these trends, we also find that the use of US-specific terminology has marginally decreased since 2005 - once again suggesting that biomedical scientists seem to have abandoned the use of race and ethnicity in exchange for geographic terminology to enact population differences in their research.

### Appendices 

Here is a list of the terms that were used to classify the OMB/US Census terms in the final two sets of analyses.

```{r, warning=FALSE, message=FALSE, echo = FALSE}
setwd("~/Documents/Diversity/Data")
terms_table <- read_csv("diversity_project - h3_dictionary.csv") 
terms_table <- terms_table %>% select(asian, black, hispanic_latinx, native_american,	pacific_islander,white)
terms_table[is.na(terms_table)] <- ""
DT::datatable(terms_table)
```


