---
title: "Hypothesis 2"
description: "We hypothesize that (a) US-based researchers are most likely to use population terms and (b) that US-specific racial and ethnic categories have grown in use in biomedical and diversity abstracts over the past three decadess."
output: html_document
weight: 3
---

To test Hypothesis 3, we evaluated the geographical basis of population terminology use and variation in US-specific racial and ethnic categories over time. First, we extracted and parsed all of the address information of authors in our diversity corpus, recoding these data into ISO Alpha-2 country codes. We then examined how often all population terms (i.e. the dictionary of 8,200+ terms) were used from researchers in each country across both the biomedical and diversity samples. 

```{r, include=FALSE}
rm(list = ls())

for (pkg in c("tidyverse", "igraph", "tidytext", "ggplot2", "DT", "maditr",
              "grid", "gridExtra", "reshape2", "extrafont", "data.table",
              "plotly", "stringi", "networkD3")) {library(pkg, character.only = TRUE)}

h2_omb_counts <- read_rds("~/git/diversity/data/text_results/h2_results/h2_all_omb_counts.rds") 
h2_omb_prcs <- read_rds("~/git/diversity/data/text_results/h2_results/h2_all_omb_prcs.rds") 

```

Now that we know US-based researchers are most likely to use population terms (a result we fully expected), we want to further examine how OMB/US Census terms are used in our two corpora. To ensure that we are capturing US-centric terminology, we [compiled a dictionary](https://growthofdiversity.netlify.app/analyses/dictionaries) used to catch to the main OMB/US Census terms, including Asian, White, Black, Native American, Pacific Islander, Hispanic/Latinx as well as more general terms like race and ethnicity. Given that these terms can also refer to populations outside of the US, we created bigrams of US-centric terms to ensure they were specific to the American context (i.e. we constructed bigrams of all adjacent uses of African and American). 

```{r figure 3c, message = FALSE, warning = FALSE, echo=FALSE}

# graphing us-specific terms over time using census categories 
# https://www.census.gov/topics/population/race/about.html

figure_3C <- h2_omb_counts %>% 
  group_by(term, year) %>% 
  summarise(n = sum(count)) %>% 
  arrange(-n) %>% 
  ggplot() + 
  geom_line(aes(y = n, x = year, colour = term), size = 1.3, stat="identity") + 
  theme_minimal() +
  labs(title = "     Figure 3C. Total Growth of OMB/US Census Terms \n in Biomedical Sample (1990-2020)",
       color =  "Term") + 
  ylab("Term Count") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(labels=c("asian","black", "diversity (social)",
                              "ethnicity", "hispanic/latinx", "native-american",
                              "pacific-islander","race","white"),
                     values=c("#005BB7", "#E86F00", "#800080", 
                              "#61B329", "#3B9AB2", "#ff8396", 
                              "#000000", "#f2be00", "#CC0000")) +
  scale_x_continuous(limits = c(1990, 2020),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020)); ggplotly() 

#                              "#E86F00", # orange 
#                              "#61B329", # light green
#                              "#f2be00", # mustard
#                              "#800080", # purple
#                              "#000000", # dark green 
#                              "#005BB7", # dark blue 
#                              "#CC0000", # red 
#                              "#ff8396", # pinkish 
#                              "#3B9AB2" # light blue 
```

Figure 3C shows the OMB US Census terms are growing over time, but this growth is quite subdued compared to our previous results. For example, no term set exceeds more than 400 uses in a given year. Of these category sets, ethnicity grows the most followed by Asian, White, race, and Black. 


```{r figure_3D, message = FALSE, warning=FALSE, echo=FALSE}
# graphing terms over time 
figure_3D <- h2_omb_prcs %>% 
  ggplot() + 
  geom_line(aes(y = prc_black, x = year, colour = "black"), size = 1.3, stat="identity") +
  geom_line(aes(y = prc_white, x = year, colour = "white"),size = 1.3, stat="identity") +
  geom_line(aes(y = prc_asian, x = year, colour = "asian"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_hispanic, x = year, colour = "hispanic/latinx"),size = 1.3, stat="identity") +
  geom_line(aes(y = prc_native_american, x = year, colour = "native-american"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_pacific_islander, x = year, colour = "pacific-islander"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_ethnic, x = year, colour = "ethnicity"), size = 1.3,stat="identity") +
  geom_line(aes(y = prc_racial, x = year, colour = "race"), size = 1.3, stat="identity") +
  geom_line(aes(y = prc_diversity, x = year, colour = "diversity (social)"), size = 1.3, stat="identity") +
  labs(title = "    Figure 3D. Proportional Growth of \n US-Specific Terms in Biomedical Sample (1990-2020)",
       color =  "Term") + 
  theme_minimal() +
  ylab("Proportion of Articles") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        #legend.position = "none",
        legend.box = "vertical", 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(labels=c("asian","black", "diversity (social)",
                              "ethnicity", "hispanic/latinx", "native-american",
                              "pacific-islander","race","white"),
                     values=c("#005BB7", "#E86F00", "#800080", 
                              "#61B329", "#3B9AB2", "#ff8396", 
                              "#000000", "#f2be00", "#CC0000")) +
  scale_x_continuous(limits = c(1990, 2020),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015, 2020)); ggplotly()
```

Similarly, Figure 3D shows that the OMB/US Census terms vary little over time. While there are some brief spikes in usage, most totals never surpass 1% of abstracts in any given year. This plot further corroborates for the notion that researchers are moving away from the use of racial and ethnic terms to describe populations in their work.

### Main Takeaway 

Overall, Figures 3A-3F support our hypothesis that *(a)* US-based researchers tend to use population terminology at disproportionately higher rates than scientists from other countries and *(b)* that terminology based on the US-Census and OMB Directive 15 have increased since 1990. However, when looking more closely at these trends, we also find that the use of US-specific terminology has marginally decreased since 2005 - once again suggesting that biomedical scientists seem to have abandoned the use of race and ethnicity in exchange for geographic terminology to enact population differences in their research.

### Appendices 

Here is a list of the terms that were used to classify the OMB/US Census terms in the final two sets of analyses.

```{r, warning=FALSE, message=FALSE, echo = FALSE}
setwd("~/git/diversity/data/dictionaries/")
terms_table <- read_csv("diversity_project - h2_dictionary.csv") 
terms_table <- terms_table %>% select(asian, black, hispanic_latinx, native_american,	pacific_islander,white)
terms_table[is.na(terms_table)] <- ""
DT::datatable(terms_table)
```


