---
title: "Hypothesis 2"
description: "The use of population terminology has increased in biomedical abstracts since 1990."
weight: 2
output: html_document
---

```{r prereqs, message = FALSE, results = FALSE, echo=FALSE}
rm(list = ls())

for (pkg in c("tidyverse", "igraph", "tidytext", "ggplot2", "DT",
              "grid", "gridExtra", "reshape2", "extrafont",
               "plotly", "stringi", "networkD3")) {library(pkg, character.only = TRUE)}

setwd("~/Documents/Diversity/Data")
# loading the .csv file 
text_data <- read_csv("biomedical_corpus.csv") %>% 
  rowid_to_column(var = "id") %>% 
  rename(authors = AU, title = TI, publication = SO, #author_keywords = DE,
         abstract = AB, references = CR, year = PY, times_cited = TC, pubmed_id = UT) %>% 
  select(id, authors, title, year, publication, abstract, references, year, times_cited, pubmed_id) %>% 
# filtering 2018 articles because they seem to be incomplete 
  filter(year != "2018")
# tokenizing the abstract data into words 
abstract_data <- text_data %>% 
  unnest_tokens(word, abstract) #%>% anti_join(stop_words)
# adding custom set of stopwords 
my_stopwords <- tibble(word = c(as.character(1:14), "[0-9]", 
                                "rights", "reserved", "copyright", "elsevier", 
                                "women.recommendations", 
                                "cs.man.ac.uk", "sexes.please", "movements.baby"))
abstract_data <- abstract_data %>% anti_join(my_stopwords); rm(my_stopwords)

# lets draw all of our strings from the diversity_dictionary
setwd("~/Documents/Diversity/Data")
divictionary <- read_csv("diversity_project - h1_categories.csv") 

# pulling our list of population
all_pop_terms <- read_csv("diversity_project - h2h3_categories.csv") 
all_pop_terms <- paste(c("\\b(?i)(zcx", all_pop_terms$term, "zxc)\\b"), collapse = "|")
```

Next, we return to the diversity corpus to test Hypothesis 2. Here, we examine whether the use of population terminology has grown in biomedical abstracts over the past three decades. To do this, we build on Panofsky and Bliss’s (2017) exploration of how research scientists enact population differences using a polyvocality of terms, including geographic classifications like continents, countries, and cardinal directions (i.e. eastern, western, etc.). Drawing from their coding schema, we developed a list of over 2,200+ terms that correspond to national, continental, directional, racial, ethnic, and other general population groupings (e.g. terms like race, ethnicity and population) that biomedical researchers use to enact population differences.

```{r figure 2a, echo=FALSE}
figure_2A_prep <- abstract_data %>% 
  mutate(pop_term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population", no = word)) %>% 
  mutate(allpop_term = ifelse(test = str_detect(string = word, 
                       pattern = all_pop_terms), 
                       yes = "all population terms", no = word)) %>% 
  mutate(re_term = ifelse(test = str_detect(string = word, 
         pattern = paste(c("\\b(?i)(zqx", na.omit(divictionary$race_ethnicity), "zqx)\\b"), collapse = "|")), 
         yes = "race/ethnicity", no = word)) %>% 
  mutate(div_term = ifelse(test = str_detect(string = word, 
         pattern = paste(c("\\b(?i)(zqx", na.omit(divictionary$diversity), "zqx)\\b"), collapse = "|")), 
         yes = "diversity", no = word)) 
# these extra columns are needed because there is overap in the categories 
# (e.g. race, ethnicity and population are all all_pop_terms)

figure_2A_step2 <- figure_2A_prep %>% 
  group_by(year) %>% 
  count(pop_term, sort = TRUE) %>% 
  filter(pop_term == "population") %>% 
  rename(term = pop_term)

figure_2A_step3 <- figure_2A_prep %>% 
  group_by(year) %>% 
  count(allpop_term, sort = TRUE) %>% 
  filter(allpop_term == "all population terms") %>% 
  rename(term = allpop_term)

figure_2A_step4 <- figure_2A_prep %>% 
  group_by(year) %>% 
  count(re_term, sort = TRUE) %>% 
  filter(re_term == "race/ethnicity") %>% 
  rename(term = re_term)

figure_2A_step5 <- figure_2A_prep %>% 
  group_by(year) %>% 
  count(div_term, sort = TRUE) %>% 
  filter(div_term == "diversity") %>% 
  rename(term = div_term)

figure_2A_data <- rbind(figure_2A_step2, figure_2A_step3, figure_2A_step4, figure_2A_step5)
rm(figure_2A_step2, figure_2A_step3, figure_2A_step4, figure_2A_step5)
   
figure_2A <- figure_2A_data %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), size = 1.3, stat="identity") + 
  theme_minimal() +
  labs(title = "     Figure 2A. Total Growth of \n Collapsed Diversity Terms (1990-2017)",
       color =  "Term") + 
  #scale_linetype_manual( values=linetype) +
  ylab("Term Count") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(values=c("#61B329", "#E21F00", "#E2B306", "#00468b")) +
  scale_x_continuous(limits = c(1990, 2017),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015)); ggplotly()
```

Text will be here.

```{r figure 2b, echo=FALSE}
# proportional graph 

all_pop_terms_trends <- figure_2A_prep %>%   
  mutate(population = ifelse(test = str_detect(string = pop_term, 
                       pattern = "population"), yes = 1, no = 0)) %>% 
  mutate(all_pop_terms = ifelse(test = str_detect(string = allpop_term, 
                       pattern = "all population terms"), yes = 1, no = 0)) %>% 
  mutate(diversity = ifelse(test = str_detect(string = div_term, 
                       pattern = "diversity"), yes = 1, no = 0)) %>% 
  mutate(race_ethnicity = ifelse(test = str_detect(string = re_term, 
                       pattern = "race/ethnicity"), yes = 1, no = 0)) 
# total articles each year 
all_pop_prc_counts <- text_data %>% 
  group_by(year) %>% 
  count(year) %>% 
  rename(total = n) 
# articles with term mentioned each year 
all_pop_prc_counts <- all_pop_terms_trends %>% 
  filter(population == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_population = n_distinct(pubmed_id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_population = round(cnt_population / total * 100, digits = 2))
all_pop_prc_counts <- all_pop_terms_trends %>% 
  filter(all_pop_terms == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_all_pop = n_distinct(pubmed_id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_all_pop = round(cnt_all_pop / total * 100, digits = 2))
all_pop_prc_counts <- all_pop_terms_trends %>% 
  filter(diversity == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_diversity = n_distinct(pubmed_id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_diversity = round(cnt_diversity / total * 100, digits = 2))
all_pop_prc_counts <- all_pop_terms_trends %>% 
  filter(race_ethnicity == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_race_ethnicity = n_distinct(pubmed_id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_race_ethnicity = round(cnt_race_ethnicity / total * 100, digits = 2))

# graphing the patterns over time
figure_2B <- all_pop_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_population, x = year, colour = "population"), size = 1.3, stat="identity") +
  geom_line(aes(y = prc_all_pop, x = year, colour = "all population terms"), size = 1.3, stat="identity") + 
  geom_line(aes(y = prc_diversity, x = year, colour = "diversity"), size = 1.3, stat="identity") + 
  geom_line(aes(y = prc_race_ethnicity, x = year, colour = "race/ethnicity"), size = 1.3, stat="identity") + 
  labs(title = "    Figure 2B. Proportional Growth of \n Collapsed Diversity Terms (1990-2017)",
       color =  "Term") + 
  theme_minimal() +
  ylab("Proportion of Articles") +
  theme(axis.title.x = element_blank(), 
        #axis.title.y = element_blank(), 
        #legend.position = "none",
        legend.box = "vertical", 
        legend.position = "none",
        legend.title = element_text(size = 10, hjust = 0.5, face="bold"),
        plot.title = element_text(size=12, hjust = 0.5)) +
  scale_color_manual(values=c("#61B329", "#E21F00", "#E2B306", "#00468b")) +
  scale_x_continuous(limits = c(1990, 2017),
                     breaks = c(1990, 1995, 2000, 2005, 2010, 2015)); ggplotly()
```

Text will be here. 

### Main Takeaway 

Text will be here.

### Appendix 

Here is a list of the 2,200+ terms that were collapsed into the "all population terms" category in these analyses. You can use the search box to find a specific term you are interested in. Please note that this dictionary is still in progress and the sub/categories are both fluid and imperfect. If you have suggestions about what to add and/or reclassify, please send an email to the authors using the email link in the top-right corner.

```{r, warning=FALSE, message=FALSE, echo = FALSE}
setwd("~/Documents/Diversity/Data")
all_pop_terms_table <- read_csv("diversity_project - h2h3_categories.csv")
all_pop_terms_table <- all_pop_terms_table #%>% select(term, category)
all_pop_terms_table[is.na(all_pop_terms_table)] <- ""
DT::datatable(all_pop_terms_table)
```
