---
title: "The Growth of Diversity"
author: "Brandon L. Kramer (UVA, Biocomplexity Institute) & Catherine Lee (Rutgers, Sociology)"
output: html_document 
---

### Project Overview

Over the course of the 25 years, there have been persistent calls from both governmental and scientific circles to include more diverse populations in biomedical research (Epstein 2008). Among the many initatives that have been put in place since, the National Institute of Health (NIH) spearheaded this process mandating that all federally funded clinical research work to include women and racial/ethinic minorities in their research, citing concerns over population-specific diseases and potential side-effects that may only impact certain groups (Oh et al. 2015). In this project, we are interested in researching how this policy and other inclusion-based initiatives have contributed to the growth of diversity-related terminiology in biomedical research. To do this, we will be using aspects of computational text analysis to outline the emergence of diversity-related terms and how they vary by some common co-variates such as geography and discipline. 

### Data Overview

In our analyses below, we are examining the growth of in the use of the term "diversity". To do this, we drew from the MEDLINE database in Web of Science, using the search terms ["TS=(diversity)" from 1990-2017 for human research only](https://docs.google.com/spreadsheets/d/1RNpPAyowuDfA5p8Sx7nEIw1a7Z8VWFoMFc3WPQ0QELw/edit#gid=933487936). This search provided 71,528 total results, which we extracted using the bibliometrix package in R (Aria and Cuccurillo 2017). Next, we converted the abstracts of these articles to a text corpus and then used tidytext - a package designed for computational text analysis in the open-source software R (Silge and Robinson 2016) - to analyze patterns with the abstracts of these data. Below is a summary of the results, but the replication code can be found on [Brandon's Github page](github.com/brandonleekramerdiversity).

```{r setup, include = FALSE}
#rm(list = ls())
#knitr::opts_knit$set(root.dir = "~/Documents/Diversity/Data")
knitr::opts_knit$set(root.dir = "~/diversity-data")
#knitr::opts_knit$set(root.dir = "C:/Users/bkram/CloudStation/Diversity/Data")
```

### Overall Growth of Publications 

In this first chunk, we load our data and examine the overall growth of articles from our search query. As we can see, there is a pretty sizable growth in scientific literature that uses the term diversity - from about 500 times in 1990 to over 5000 in 2017.  

```{r load packages, message = FALSE, results = FALSE, echo=FALSE}
# loading packages 
for (pkg in c("tidyverse", "igraph", "tidytext", "ggplot2", "tm",
               "data.table", "countrycode",
              "plotly", "stringi", "networkD3")) {library(pkg, character.only = TRUE)}
# loading the .csv file 
text_data <- read_csv("diversity_corpus.csv") %>% 
# filtering 2018 articles because they seem to be incomplete 
  filter(year != "2018") 
```

```{r wos study growth, fig.width=9.5, echo=FALSE}
# checking to see how the overall data looks 
pubs_by_year <- text_data %>% 
  group_by(year) %>% count(year, sort = TRUE) %>% ungroup() %>% 
  ggplot() + geom_line(aes(y = n, x = year), stat="identity") + 
  labs(title = "Figure 1. Growth in Diversity-Related Publications from 1990-2017",
       caption = "Data Source: Web of Science") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()); ggplotly()
```

### Word Frequencies Over Time

Next, we want to look at word frequencies by year in the literature. Note that we did remove some frequently occurring words that are not really relevant to our dataset (i.e. custom stopwords), but these do not systematically alter our results. 

```{r text analysis of abstracts, messages = FALSE, warning=FALSE, echo=FALSE}
# tokenizing the abstract data into words 
abstract_data <- text_data %>% 
  unnest_tokens(word, abstract) %>% 
  anti_join(stop_words)
# adding custom set of stopwords 
my_stopwords <- tibble(word = c(as.character(1:14),  "[0-9]", 
                                "rights", "reserved", "copyright", "elsevier"))
abstract_data <- abstract_data %>% anti_join(my_stopwords); rm(my_stopwords)
# most frequent word count in abstracts 
abstract_data %>%
  count(word, sort = TRUE)
```

```{r abstract word count, messages = FALSE, warning=FALSE, include=FALSE}
# looking at word frequencies by year 
abstract_data %>%
  group_by(year) %>% 
  count(word, sort = TRUE) %>% ungroup()
```

### Growth in General Population Terms Over Time

Now that we have the basics out of the way, we want to look at how the most relevant words vary over time. We chose to include words like diversity and its various sub-types (i.e. genetic, cultural, and population) as well as racially-specific and ethnically-specific terms. 

```{r figure 2a, fig.width=9.5, echo=FALSE}
# recoding race*, ethnicity* and cultural* terminology
general_pop_terms <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(race|racial|racially|ethnic|ethnicity|ethnically)\\b"), 
                       yes = "race/ethnicity", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, pattern = "\\b(?i)(man|men|male|males|boy|boys|woman|women|female|females|girl|girls|XX|XY|XXY|sex|sexes|gender|genders|sexed|gendered|transgender|transgendered|transsexual|transsexuals|maletofemale|femaletomale)\\b"), yes = "sex/gender", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, pattern = "\\b(?i)(sexuality|sexualities|homosexual|homosexuals|homosexuality|heterosexual|heterosexuals|heterosexuality|bisexual|bisexuality|monogamous|monogamy|polygamous|polygamy|polygyny|polygynous|lgb|lgbt|lgbtq|lesbian|lesbians|gay|gays|queer|queers)\\b"), 
                       yes = "sexuality", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(culture|cultural|culturally)\\b"), 
                       yes = "cultural", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern =
  "\\b(?i)(ancestry|ancestries|ancestral|ancestor|ancestors|descent|descendent|descendents)\\b"), 
                       yes = "ancestry", no = term)) %>%
# grouping by year and filter relevant population terms 
  group_by(year) %>% count(term, sort = TRUE) %>% ungroup() %>% 
  filter(term == "diversity" | term == "cultural" | term == "genetic" | 
         term == "sex/gender" | term == "sexuality" |  
         term == "population" | term == "race/ethnicity" | term == "ancestry") %>% 
# graphing terms over time 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), stat="identity") + 
  labs(title = "Figure 2A. Total Growth in Diversity-Related Terminology (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

As the graph shows, the rise of diversity does not necessarily mean that researchers focus on race or ethnicity. This could mean that diversity is being used as a catch-all in the scientific literature (i.e. that the multiplicity of the term makes it mean anything and everything) or that diversity is most used in fields like immunology or oncology. We will explore that hypothesis a bit more below. Let's take a look at our first graph.

```{r figure 2b, fig.width=9.5, echo=FALSE}
# recoding race*, ethnicity* and cultural* terminology
general_pop_terms <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(race|racial|racially|ethnic|ethnicity|ethnically)\\b"), 
                       yes = "race/ethnicity", no = word)) %>% 
  mutate(di_racial = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(race/ethnicity)\\b"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(culture|cultural|culturally)\\b"), 
                       yes = "cultural", no = term)) %>% 
  mutate(di_cultural = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(cultural)\\b"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population", no = term)) %>% 
  mutate(di_population = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(population)\\b"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern =
  "\\b(?i)(ancestry|ancestries|ancestral|ancestor|ancestors|descent|descendent|descendents)\\b"), 
                       yes = "ancestry", no = term)) %>% 
  mutate(di_ancestry = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(ancestry)\\b"), yes = 1, no = 0)) %>% 
  mutate(di_diversity = ifelse(test = str_detect(string = term, 
                       pattern = "diversity"), yes = 1, no = 0)) %>% 
  mutate(di_genetic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(genetic)\\b"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, pattern = "\\b(?i)(man|men|male|males|boy|boys|woman|women|female|females|girl|girls|XX|XY|XXY|sex|sexes|gender|genders|sexed|gendered|transgender|transgendered|transsexual|transsexuals|maletofemale|femaletomale)\\b"), yes = "sex/gender", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, pattern = "\\b(?i)(sexuality|sexualities|homosexual|homosexuals|homosexuality|heterosexual|heterosexuals|heterosexuality|bisexual|bisexuality|monogamous|monogamy|polygamous|polygamy|lgb|lgbt|lgbtq|lesbian|lesbians|gay|gays|queer|queers)\\b"), yes = "sexuality", no = term)) %>%
  mutate(di_sexgender = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(sex/gender)\\b"), yes = 1, no = 0)) %>%
  mutate(di_sexuality = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(sexuality)\\b"), yes = 1, no = 0)) 
  

# total articles each year 
gen_pop_prc_counts <- text_data %>% 
  group_by(year) %>% 
  count(year) %>% 
  ungroup() %>% 
  rename(total = n) 
# articles with term mentioned each year 
gen_pop_prc_counts <- general_pop_terms %>% filter(di_racial == 1) %>% 
  group_by(year) %>% summarise(cnt_racial = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>%  
  mutate(prc_racial = round(cnt_racial / total * 100, digits = 2))
#gen_pop_prc_counts <- general_pop_terms %>% filter(di_ethnic == 1) %>% 
#  group_by(year) %>% summarise(cnt_ethnic = n_distinct(id)) %>% 
#  right_join(gen_pop_prc_counts, by = "year") %>% 
#  mutate(prc_ethnic = round(cnt_ethnic / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_cultural == 1) %>% 
  group_by(year) %>% summarise(cnt_cultural = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_cultural = round(cnt_cultural / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_population == 1) %>% 
  group_by(year) %>% summarise(cnt_population = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_population = round(cnt_population / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_ancestry == 1) %>% 
  group_by(year) %>% summarise(cnt_ancestry = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_ancestry = round(cnt_ancestry / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_diversity == 1) %>% 
  group_by(year) %>% summarise(cnt_diversity = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_diversity = round(cnt_diversity / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_genetic == 1) %>% 
  group_by(year) %>% summarise(cnt_genetic = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_genetic = round(cnt_genetic / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_sexgender == 1) %>% 
  group_by(year) %>% summarise(cnt_sexgender = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_sexgender = round(cnt_sexgender / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_sexuality == 1) %>% 
  group_by(year) %>% summarise(cnt_sexuality = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_sexuality = round(cnt_sexuality / total * 100, digits = 2))

# graphing terms over time 
gen_pop_prc_counts <- gen_pop_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_ancestry, x = year, colour = "ancestry"), stat="identity") +
  geom_line(aes(y = prc_cultural, x = year, colour = "cultural"), stat="identity") +
  geom_line(aes(y = prc_diversity, x = year, colour = "diversity"), stat="identity") +
  geom_line(aes(y = prc_racial, x = year, colour = "race/ethnicity"), stat="identity") +
  geom_line(aes(y = prc_genetic, x = year, colour = "genetic"), stat="identity") +
  geom_line(aes(y = prc_sexgender, x = year, colour = "sex/gender"), stat="identity") +
  geom_line(aes(y = prc_sexuality, x = year, colour = "sexuality"), stat="identity") +
  geom_line(aes(y = prc_population, x = year, colour = "population"), stat="identity") +
  labs(title = "Figure 2B. Proportional Growth in Diversity-Related Terminology (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly() 
```

While Figure 2A looks at the raw counts of words mentioned across our dataset, we also want to know what percentage of abstracts that these terms are mentioned in. Figure 2B shows how often general diversity-related terms occur within the corpus of abstracts. As you can see, "diversity" still occurs most often, which is not surprising given our search terms bias the inclusion of that term in a big way. (Note that the search terms include any entry that has diversity in its title, abstract or keywords, which is why we don't see diversity included in 100% of our dataset in any given year.) 

What is interesting about that term is that "diversity", as a term that arises in a certain percentage of scientfic abstracts, actually falls through the early 1990s before rebounding and consistently rising thereafter. Note that the NIH's Health Revitalization Act was instituted in 1993, which may have contributed to this rise of diversity-related terminology. Population, genetic and cultural follow similar upward trends over time, though the rise is much more subdued. On the other hand, race, ancestry and ethnicity all increase as a percentage until the mid-2000's before declining slightly thereafter. Our analyses do not address why these declines occur, though some of the results below suggest that the use of continental and national terms overtake the use of race and ethnicity after 2003, which also happens to be around the time the Human Genome Project was completed. 

### Growth in US-Specific Population Terms Over Time

Now we want to look at all [US-specific population terminology](https://www.census.gov/topics/population/race/about.html). This is not much different than what we have already done except for the small technical problem of tidytext separating hyphenated words in our original analysis (e.g. African-Americans). While the package quanteda has an option to unnest tokens while ignoring hyphenated words, quanteda takes a lot of RAM and ushers in other potential methodological inconsistencies of switching back and forth between two packages. Instead, we found bigrams of key US-specific terms in [tidytext](https://www.tidytextmining.com/ngrams.html) and then add them back into our larger dataset for trend anaylsis. 

```{r figure 3a, fig.width=9.5, echo=FALSE}
# recoding frequently occuring population-related bigrams 
us_pop_bigrams <- text_data %>% 
  unnest_tokens(bigram, abstract, token = "ngrams", n = 2) %>% 
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(african american|african americans|afro american|afro americans)\\b"), 
                       yes = "african-american", no = bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(native american|native americans|american indian|american indians|alaskan native|alaskan natives|alaska native|alaska natives)\\b"), 
                       yes = "native-american", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(pacific islander|pacific islanders|native hawaiian|native hawaiians)\\b"), 
                       yes = "pacific-islander", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(mexican american|mexican americans)\\b"), 
                       yes = "hispanic-american", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(cambodian american|cambodian americans|chinese american|chinese americans|indian american|indian americans|japanese american|japanese americans|korean american|korean americans|malaysian american|malaysian americans|pakistani american|pakistani americans|filipino american|filipino americans|thai american|thai americans|vietnamese american|vietnamese americans)\\b"), 
  yes = "asian-american", no = rc_bigram)) %>%
  select(-bigram) %>% 
  filter(rc_bigram == "african-american" | 
         rc_bigram == "native-american" | 
         rc_bigram == "hispanic-american" |
         rc_bigram == "asian-american" |
         rc_bigram == "pacific-islander") %>% 
  count(rc_bigram, year, sort = TRUE) %>% 
  rename(word = rc_bigram)

# combining the hyphenated words to our full dataset 
us_pop_counts <- abstract_data %>% 
  group_by(year) %>% 
  count(word, sort = TRUE) %>% 
  ungroup() %>% 
  select(word, year, n) %>% 
  bind_rows(us_pop_bigrams) %>% 
  arrange(-n)

# graphing us-specific terms over time using census categories 
# https://www.census.gov/topics/population/race/about.html
us_pop_terms <- us_pop_counts %>% 
  mutate(term = ifelse(test = str_detect(string = word,
                       pattern = "\\b(?i)(black|blacks|african-american)\\b"), 
                       yes = "black", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(white|whites|caucasian|caucasians)\\b"), 
                       yes = "white", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(asian|asians|asian-american)\\b"), 
                       yes = "asian", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
  pattern = "\\b(?i)(hispanic|hispanic-american|latino|latinos|latina|latinos|latinx|latinxs)\\b"), 
                       yes = "hispanic/latinx", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(race|racial|racially)\\b"), 
                       yes = "race", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(ethnic|ethnicity|ethnically)\\b"), 
                       yes = "ethnicity", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(native-american)\\b"), 
                       yes = "native-american", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(pacific-islander)\\b"), 
                       yes = "pacific-islander", no = term)) %>%
  filter(term == "black" | term == "white" | term == "asian" |
         term == "hispanic/latinx" | term == "race" | term == "ethnicity" |
         term == "native-american" | term == "pacific-islander" ) %>%  
  group_by(term, year) %>% summarise(n = sum(n)) %>% arrange(-n) %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), stat="identity") + 
  labs(title = "Figure 3A. Total Growth in US-Specific Terminology (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

To make this graph of all US-specific terminology, we recoded all hyphenated words that we consider US-specific population terms, including African-American, Mexican-American, American-Indian, Native American, Alaskan Indian, and Pacific Islander. We recoded these variations into the five main racial/ethnic categories used on the US Census (White, Black, Hispanic, Native American, and Pacific Islander) and included them alongside all the variations of race and ethnicity. We also included our baseline term diversity to help compare the growth of US-specific terms to our other analyses. As we can see here, there is not a ton of growth in US-specific diversity-related terminology over time (at least in comparison with the term diversity). For those that the growth of diversity is largely related to an increased proclivity for scientists to study race/ethnicity in accordance with a US-specific research program, this graph essentially shows that this hypothesis is pretty weakly supported. 

```{r figure 3b, fig.width=9.5, echo=FALSE}

# recoding frequently occuring population-related bigrams 
us_pop_bigrams <- text_data %>% 
  unnest_tokens(bigram, abstract, token = "ngrams", n = 2) %>% 
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(african american|african americans|afro american|afro americans)\\b"), 
                       yes = "african-american", no = bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(native american|native americans|american indian|american indians|alaskan native|alaskan natives|alaska native|alaska natives)\\b"), 
                       yes = "native-american", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(pacific islander|pacific islanders|native hawaiian|native hawaiians)\\b"), 
                       yes = "pacific-islander", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(mexican american|mexican americans)\\b"), 
                       yes = "hispanic-american", no = rc_bigram)) %>%
  mutate(rc_bigram = ifelse(test = str_detect(string = bigram,
  pattern = "\\b(?i)(cambodian american|cambodian americans|chinese american|chinese americans|indian american|indian americans|japanese american|japanese americans|korean american|korean americans|malaysian american|malaysian americans|pakistani american|pakistani americans|filipino american|filipino americans|thai american|thai americans|vietnamese american|vietnamese americans)\\b"), 
  yes = "asian-american", no = rc_bigram)) %>%
  select(-bigram) %>% 
  filter(rc_bigram == "african-american" | 
         rc_bigram == "native-american" | 
         rc_bigram == "hispanic-american" |
         rc_bigram == "asian-american" |
         rc_bigram == "pacific-islander") %>% 
  count(rc_bigram, year, sort = TRUE) %>% 
  rename(word = rc_bigram)

# combining the hyphenated words to our full dataset 
us_pop_counts <- abstract_data %>% 
  group_by(year) %>% 
  count(word, pubmed_id, sort = TRUE) %>% 
  ungroup() %>% 
  select(word, pubmed_id, year, n) %>% 
  bind_rows(us_pop_bigrams) 

# graphing us-specific terms over time 
us_pop_counts <- us_pop_counts %>% 
  mutate(term = ifelse(test = str_detect(string = word,
                       pattern = "\\b(?i)(black|blacks|african-american)\\b"), 
                       yes = "black", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(white|whites|caucasian|caucasians)\\b"), 
                       yes = "white", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(asian|asians|asian-american)\\b"), 
                       yes = "asian", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
  pattern = "\\b(?i)(hispanic|hispanic-american|latino|latinos|latina|latinos|latinx|latinxs)\\b"), 
                       yes = "hispanic/latinx", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(race|racial|racially)\\b"), 
                       yes = "race", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(ethnic|ethnicity|ethnically)\\b"), 
                       yes = "ethnicity", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(native-american)\\b"), 
                       yes = "native-american", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(pacific-islander)\\b"), 
                       yes = "pacific-islander", no = term)) %>% 
  
  mutate(black = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(black)\\b"), yes = 1, no = 0)) %>% 
  mutate(white = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(white)\\b"), yes = 1, no = 0)) %>%
  mutate(asian = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(asian)\\b"), yes = 1, no = 0)) %>%
  mutate(hispanic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(hispanic/latinx)\\b"), yes = 1, no = 0)) %>% 
  mutate(native_american = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(native-american)\\b"), yes = 1, no = 0)) %>% 
  mutate(pacific_islander = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(pacific-islander)\\b"), yes = 1, no = 0)) %>% 
  mutate(racial = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(race)\\b"), yes = 1, no = 0)) %>% 
  mutate(ethnic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(ethnicity)\\b"), yes = 1, no = 0)) 

# total articles each year 
us_pop_prc_counts <- text_data %>% 
  group_by(year) %>% 
  count(year) %>% 
  ungroup() %>% 
  rename(total = n) 

# articles with term mentioned each year 
us_pop_prc_counts <- us_pop_counts %>% 
  filter(black == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_black = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_black = round(cnt_black / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% 
  filter(white == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_white = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_white = round(cnt_white / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% 
  filter(asian == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_asian = n_distinct(pubmed_id)) %>% 
  left_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_asian = round(cnt_asian / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% 
  filter(hispanic == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_hispanic = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_hispanic = round(cnt_hispanic / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% 
  filter(native_american == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_native_american = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_native_american = round(cnt_native_american / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% 
  filter(pacific_islander == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_pacific_islander = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_pacific_islander = round(cnt_pacific_islander / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% 
  filter(racial == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_racial = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_racial = round(cnt_racial / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% filter(ethnic == 1) %>% 
  group_by(year) %>% 
  summarise(cnt_ethnic = n_distinct(pubmed_id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_ethnic = round(cnt_ethnic / total * 100, digits = 2))

# graphing terms over time 
us_pop_prc_counts_graph <- us_pop_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_black, x = year, colour = "black"), stat="identity") +
  geom_line(aes(y = prc_white, x = year, colour = "white"), stat="identity") +
  geom_line(aes(y = prc_asian, x = year, colour = "asian"), stat="identity") +
  geom_line(aes(y = prc_hispanic, x = year, colour = "hispanic/latinx"), stat="identity") +
  geom_line(aes(y = prc_native_american, x = year, colour = "native-american"), stat="identity") +
  geom_line(aes(y = prc_pacific_islander, x = year, colour = "pacific-islander"), stat="identity") +
  geom_line(aes(y = prc_ethnic, x = year, colour = "ethnicity"), stat="identity") +
  geom_line(aes(y = prc_racial, x = year, colour = "race"), stat="identity") +
  labs(title = "Figure 3B. Proportional Growth in US-Specific Diversity Terminology (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly() 
```

When we look at those trends as terms that occur with in a percentage of articles each year, the hypothesis that race/ethnicity is growing in use is even less viable. We can see that US-specific population terminology increases throughout the 1990s and early-2000s before declining after that. As I mentioned above, it is unclear what exactly might be causing this, but one possibility is that national and continental terms replace US-specific research as international collaboration increased after the completion of the Human Genome Project. 

### The Multiplicity of Population Terminology

While that graph was quite informative, we do not see much variation in how race/ethinicity vary over time in comparison with terms like diversity. One possibility is that variation in population terms is "diluted" by both the multiplicity of population terminology and the ambiguity of how those terms are used (M'Charek 2013; Panofsky and Bliss 2017). Thus, the next logical step is to collapse all possible population terms into one category and graph that over time. Both Lee (2009) and Kramer (2019) as well as others (e.g. Panofsky and Bliss 2017) have found that biomedical researchers continue to use population-specific terminology to reinforce the notion of population differences in various biological markers. Below, we have collapsed several population-specific terms into one category based. [This list of terms](https://docs.google.com/spreadsheets/d/1Gs-aSgXlnwo8rP3wnpMB0kdrU_VS76aDT3DCiI8OmVU/edit#gid=0) which were developed out of Kramer's (2019) dissertation work. While we won't claim that this is an exhaustive list of all the population terms that exist in the world, it is a fairly comprehensive list of over 2,100 terms (also see below in text networks section). Let's take a look how much the use of all these population-specific terms grew over time.

```{r collapsing population specific terms, message = FALSE, results = FALSE, warning = FALSE, echo=FALSE}
# pulling our list of population
all_pop_terms <- read_csv("population_terms.csv") 
all_pop_terms <- paste(c("\\b(?i)(zcx", all_pop_terms$term, "zxc)\\b"), collapse = "|")
```

```{r figure 4a, fig.width=9.5, echo=FALSE}
all_pop_terms_graph <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = all_pop_terms), 
                       yes = "all population terms", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(race|racial|racially|ethnic|ethnicity|ethnically)\\b"), 
                       yes = "race/ethnicity", no = term)) %>% 
  group_by(year) %>% count(term, sort = TRUE) %>% ungroup() %>% 
  filter(term == "diversity" | term == "race/ethnicity" | 
         term == "population" | term == "all population terms") %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), stat="identity") + 
  labs(title = "Figure 4A. Total Growth in Diversity-Related Terminology (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

Interesting! After we collapsed all of the population terms together into one category, we see a dramatic growth that goes well beyond even the term "diversity" that was so high to begin with. This give some support to the notion that multiplicity and ambiguity could be mechanisms that drive the growth of diversity-related terms. More specifically, following M'Charek (2013), we are arguing that race is a multiplicity; it is a discourse that can link to *various* material referents depending on the context it is embedded within. The definition of race can be dependent on the technologies or researchers that are enacting the concept to carry out specific ends. Because of its multiplicitous nature, race also becomes ambiguous and uncertain. Race can be many things, which obscures the precision of any specific claims that can be made about it. Panofsky and Bliss (2017) find that population terms grow in use over time in the context of biomedical studies, but there use is often shifting and imprecise. The authors argue that this ambiguity seems to even reinforce authority in some scientific communities. Building on that work, we do see evidence that variability in population terms at multiple levels does increase over time. We see the same trends when we look the growth of all population claims as a percentage as well (below).

```{r figure 4b, fig.width=9.5, echo=FALSE}
all_pop_terms_trends <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population*", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = all_pop_terms), 
                       yes = "all population terms", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(race|racial|racially|ethnic|ethnicity|ethnically)\\b"), 
                       yes = "race/ethnicity", no = term)) %>% 
  mutate(population = ifelse(test = str_detect(string = term, 
                       pattern = "\\b((?<!all )population)\\b"), yes = 1, no = 0)) %>% 
  mutate(all_pop_terms = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(all population terms)\\b"), yes = 1, no = 0)) %>% 
  mutate(diversity = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(diversity)\\b"), yes = 1, no = 0)) %>% 
  mutate(race_ethnicity = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(race/ethnicity)\\b"), yes = 1, no = 0)) 
# total articles each year 
all_pop_prc_counts <- text_data %>% group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n) 
# articles with term mentioned each year 
all_pop_prc_counts <- all_pop_terms_trends %>% filter(population == 1) %>% 
  group_by(year) %>% summarise(cnt_population = n_distinct(id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_population = round(cnt_population / total * 100, digits = 2))
all_pop_prc_counts <- all_pop_terms_trends %>% filter(all_pop_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_all_pop = n_distinct(id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_all_pop = round(cnt_all_pop / total * 100, digits = 2))
all_pop_prc_counts <- all_pop_terms_trends %>% filter(diversity == 1) %>% 
  group_by(year) %>% summarise(cnt_diversity = n_distinct(id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_diversity = round(cnt_diversity / total * 100, digits = 2))
all_pop_prc_counts <- all_pop_terms_trends %>% filter(race_ethnicity == 1) %>% 
  group_by(year) %>% summarise(cnt_race_ethnicity = n_distinct(id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_race_ethnicity = round(cnt_race_ethnicity / total * 100, digits = 2))

# graphing the patterns over time
all_pop_prc_counts <- all_pop_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_population, x = year, colour = "population"), stat="identity") +
  geom_line(aes(y = prc_all_pop, x = year, colour = "all population terms"), stat="identity") + 
  geom_line(aes(y = prc_diversity, x = year, colour = "diversity"), stat="identity") + 
  geom_line(aes(y = prc_race_ethnicity, x = year, colour = "race/ethnicity"), stat="identity") + 
  labs(title = "Figure 4B. Proportional Growth in Collapsed Diversity-Related Terminology (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

Now, let's see what the growth of all these population terms when we break them into general, national, continental, and ethnic-specific variations over time. Here, general refers to terms like race, ethnicity, and other terms related broadly to racial terms. National and continental terms are self-explanatory. Ethnic-specific terms refers to all possible ethnic terms we could find in the course of generating these terms. That said, we still have some work to do on refining these various terms before publication.

```{r breaking apart collapsed population terms, message = FALSE, warning = FALSE, echo=FALSE}
# pulling in concatenated population terms from our dataset 
general_pop_terms <- read_csv("population_terms.csv") %>% 
  filter(sub_category == "population_general")
us_specific_terms <- read_csv("population_terms.csv") %>% 
  filter(sub_category == "us_specific")
continental_terms <- read_csv("population_terms.csv") %>% 
  filter(category == "continental" | category == "subcontinental")
ling_religious_terms <- read_csv("population_terms.csv") %>% 
  filter(category == "linguistic_religious")
national_terms <- read_csv("population_terms.csv") %>% 
  filter(category == "national")
south_american_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "south_america")
african_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "africa")
north_american_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "north_america")
european_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "europe")
asian_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "asia")
all_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "south_america" | category == "africa" | category == "asia" | 
         category == "north_america" | category == "europe")
# creating detection string patterns for all the categories
general_pop_terms <- paste(c("\\b(?i)(zxz", 
                     general_pop_terms$term, "zxz)\\b"), collapse = "|")
us_specific_terms <- paste(c("\\b(?i)(zxz", 
                     us_specific_terms$term, "zxz)\\b"), collapse = "|")
continental_terms <- paste(c("\\b(?i)(zxz", 
                     continental_terms$term, "zxz)\\b"), collapse = "|")
ling_religious_terms <- paste(c("\\b(?i)(zxz", 
                     ling_religious_terms$term, "zxz)\\b"), collapse = "|")
national_terms <- paste(c("\\b(?i)(zxz", 
                     national_terms$term, "zxz)\\b"), collapse = "|")
south_american_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     south_american_ethnic_groups$term, "zxz)\\b"), collapse = "|")
african_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     african_ethnic_groups$term, "zxz)\\b"), collapse = "|")
north_american_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     north_american_ethnic_groups$term, "zxz)\\b"), collapse = "|")
european_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     european_ethnic_groups$term, "zxz)\\b"), collapse = "|")
asian_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     asian_ethnic_groups$term, "zxz)\\b"), collapse = "|")
all_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     all_ethnic_groups$term, "zxz)\\b"), collapse = "|")
```

```{r figure 5a, fig.width=9.5, echo=FALSE}
# graphing the population terms into various categories 
pop_terms_bycats_pt1 <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = general_pop_terms), 
                       yes = "general population terms", no = word)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = us_specific_terms), 
                       yes = "us-specific terms", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = continental_terms), 
                       yes = "continental terms", no = term)) %>% 
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = ling_religious_terms), 
  #                     yes = "linguistic & religious terms", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = national_terms), 
                       yes = "national terms", no = term)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = south_american_ethnic_groups), 
  #                     yes = "south american ethnic groups", no = term)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = african_ethnic_groups), 
  #                     yes = "african ethnic groups", no = term)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = north_american_ethnic_groups), 
  #                     yes = "north american ethnic groups", no = term)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = european_ethnic_groups), 
  #                     yes = "european ethnic groups", no = term)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = asian_ethnic_groups), 
  #                     yes = "asian ethnic groups", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = all_ethnic_groups), 
                       yes = "all ethnic groups", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population", no = term)) %>% 
  group_by(year) %>% count(term, sort = TRUE) %>% ungroup() %>% 
  filter(term == "diversity" | term == "population" | 
         term == "general population terms" | term == "us-specific terms" |
         term == "continental terms" | term == "national terms" | term == "all ethnic terms"
         #term == "south american ethnic groups" | term == "linguistic & religious terms" |
         #term == "african ethnic groups" | term == "north american ethnic groups" | 
         #term == "european ethnic groups" | term == "asian ethnic groups" 
         ) %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), stat="identity") + 
  labs(title = "Figure 5A. Growth in Diversity-Related Terminology by Category (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

As we can see here, the majority of growth in population-specific terms is because of an increased proclivity to mention nationalities (e.g. Italian, Japanese, Canadian, etc.), continental terms (e.g. Africa, Asia, Europe, or North and South America), and general population terms (e.g. race, ethnicity, Caucasian, African, Asian, etc.). Relative to these other general terms, we do not really see much growth in ethnic groups in any given continent or when we lump all of the different ethnic groups into one category. 

```{r figure 5b, fig.width=9.5, echo=FALSE}
# graphing the population terms into various categories 
pop_terms_bycats <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = general_pop_terms), 
                       yes = "general population terms", no = word)) %>%
  mutate(general_terms = ifelse(test = str_detect(string = term, 
                       pattern = "general population terms"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = us_specific_terms), 
                       yes = "us-specific terms", no = term)) %>%
  mutate(us_terms = ifelse(test = str_detect(string = term, 
                       pattern = "us-specific terms"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = continental_terms), 
                       yes = "continental terms", no = term)) %>%
  mutate(continental_terms = ifelse(test = str_detect(string = term, 
                       pattern = "continental terms"), yes = 1, no = 0)) %>% 
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = ling_religious_terms), 
  #                     yes = "linguistic & religious terms", no = term)) %>%
  #mutate(ling_rel_terms = ifelse(test = str_detect(string = term, 
  #                     pattern = "linguistic & religious terms"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = national_terms), 
                       yes = "national terms", no = term)) %>%
  mutate(national_terms = ifelse(test = str_detect(string = term, 
                       pattern = "national terms"), yes = 1, no = 0)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = south_american_ethnic_groups), 
  #                     yes = "south american ethnic groups", no = term)) %>%
  #mutate(south_american_terms = ifelse(test = str_detect(string = term, 
  #                     pattern = "south american ethnic groups"), yes = 1, no = 0)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = african_ethnic_groups), 
  #                     yes = "african ethnic groups", no = term)) %>%
  #mutate(african_terms = ifelse(test = str_detect(string = term, 
  #                     pattern = "african ethnic groups"), yes = 1, no = 0)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = north_american_ethnic_groups), 
  #                     yes = "north american ethnic groups", no = term)) %>%
  #mutate(north_american_terms = ifelse(test = str_detect(string = term, 
  #                     pattern = "north american ethnic groups"), yes = 1, no = 0)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = european_ethnic_groups), 
  #                     yes = "european ethnic groups", no = term)) %>%
  #mutate(european_terms = ifelse(test = str_detect(string = term, 
  #                     pattern = "european ethnic groups"), yes = 1, no = 0)) %>%
  #mutate(term = ifelse(test = str_detect(string = word, 
  #                     pattern = asian_ethnic_groups), 
  #                     yes = "asian ethnic groups", no = term)) %>% 
  #mutate(asian_terms = ifelse(test = str_detect(string = term, 
  #                     pattern = "asian ethnic groups"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = all_ethnic_groups), 
                       yes = "all ethnic groups", no = term)) %>% 
  mutate(all_ethnic_terms = ifelse(test = str_detect(string = term, 
                       pattern = "all ethnic groups"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population", no = term)) %>% 
  mutate(population = ifelse(test = str_detect(string = term, 
                       pattern = "\\b((?<!general )population)\\b"), yes = 1, no = 0)) 

# total articles each year 
bycats_prc_counts <- text_data %>% group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n) 
# articles with term mentioned each year 
bycats_prc_counts <- pop_terms_bycats %>% filter(general_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_general_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_general_terms = round(cnt_general_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(us_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_us_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_us_terms = round(cnt_us_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(continental_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_continental_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_continental_terms = round(cnt_continental_terms / total * 100, digits = 2))
#bycats_prc_counts <- pop_terms_bycats %>% filter(ling_rel_terms == 1) %>% 
#  group_by(year) %>% summarise(cnt_ling_rel_terms = n_distinct(id)) %>% 
#  right_join(bycats_prc_counts, by = "year") %>% 
#  mutate(prc_ling_rel_terms = round(cnt_ling_rel_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(national_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_national_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_national_terms = round(cnt_national_terms / total * 100, digits = 2))
#bycats_prc_counts <- pop_terms_bycats %>% filter(south_american_terms == 1) %>% 
#  group_by(year) %>% summarise(cnt_south_american_terms = n_distinct(id)) %>% 
#  right_join(bycats_prc_counts, by = "year") %>% 
#  mutate(prc_south_american_terms = round(cnt_south_american_terms / total * 100, digits = 2))
#bycats_prc_counts <- pop_terms_bycats %>% filter(african_terms == 1) %>% 
#  group_by(year) %>% summarise(cnt_african_terms = n_distinct(id)) %>% 
#  right_join(bycats_prc_counts, by = "year") %>% 
#  mutate(prc_african_terms = round(cnt_african_terms / total * 100, digits = 2))
#bycats_prc_counts <- pop_terms_bycats %>% filter(north_american_terms == 1) %>% 
#  group_by(year) %>% summarise(cnt_north_american_terms = n_distinct(id)) %>% 
#  right_join(bycats_prc_counts, by = "year") %>% 
#  mutate(prc_north_american_terms = round(cnt_north_american_terms / total * 100, digits = 2))
#bycats_prc_counts <- pop_terms_bycats %>% filter(european_terms == 1) %>% 
#  group_by(year) %>% summarise(cnt_european_terms = n_distinct(id)) %>% 
#  right_join(bycats_prc_counts, by = "year") %>% 
#  mutate(prc_european_terms = round(cnt_european_terms / total * 100, digits = 2))
#bycats_prc_counts <- pop_terms_bycats %>% filter(asian_terms == 1) %>% 
#  group_by(year) %>% summarise(cnt_asian_terms = n_distinct(id)) %>% 
#  right_join(bycats_prc_counts, by = "year") %>% 
#  mutate(prc_asian_terms = round(cnt_asian_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(all_ethnic_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_all_ethnic_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_all_ethnic_terms = round(cnt_all_ethnic_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(population == 1) %>% 
  group_by(year) %>% summarise(cnt_population = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_population = round(cnt_population / total * 100, digits = 2))

bycats_prc_counts <- bycats_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_general_terms, x = year, colour = "general population terms"), stat="identity") +
  geom_line(aes(y = prc_us_terms, x = year, colour = "us-specific terms"), stat="identity") +
  geom_line(aes(y = prc_continental_terms, x = year, colour = "continental terms"), stat="identity") +
  #geom_line(aes(y = prc_ling_rel_terms, x = year, colour = "linguistic & religious terms"), stat="identity") +
  geom_line(aes(y = prc_national_terms, x = year, colour = "national terms"), stat="identity") +
  #geom_line(aes(y = prc_south_american_terms, x = year, colour = "south american terms"), stat="identity") +
  #geom_line(aes(y = prc_african_terms, x = year, colour = "african terms"), stat="identity") +
  #geom_line(aes(y = prc_north_american_terms, x = year, colour = "north american terms"), stat="identity") +
  #geom_line(aes(y = prc_european_terms, x = year, colour = "european terms"), stat="identity") +
  #geom_line(aes(y = prc_asian_terms, x = year, colour = "asian terms"), stat="identity") +
  geom_line(aes(y = prc_all_ethnic_terms, x = year, colour = "all ethnic terms"), stat="identity") +
  geom_line(aes(y = prc_population, x = year, colour = "population"), stat="identity") +
  labs(title = "Figure 5B. Proportional Growth in Diversity-Related Terminology (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

When we look at the trends by percentage, the same national, continental and general population terminology is still most likely to grow over time. The main difference is that as a percentage, these terms taper off after the mid-2000's again, which is also the case with specific continental and ethnic terminology as well as US-specific vernacular. The exception to this plateau effect are national terms, which grow strongly all the way up to the present day. 

### Examining Diversity Across Different Geographies

The next step is to look at how the concept of "diversity" is used across the world. As the graph below demonstrates, the rise of "diversity" seems mostly to grow in the context of predominantly white, Westernized countries like the US, England, the Netherlands, Germany and Switzerland. 

```{r, message = FALSE, results = FALSE, warning = FALSE, echo=FALSE}

# pulling our list of population
all_pop_terms <- read_csv("population_terms.csv") 
all_pop_terms <- paste(c("\\b(?i)(zcx", all_pop_terms$term, "zxc)\\b"), collapse = "|")

# first we have to recode all the strings that could arise in the location colunmn as country codes 
af	<- "\\b(?i)(Afghanistan)\\b"
ax	<- "\\b(?i)(land Islands|Aland Islands)\\b"
al	<- "\\b(?i)(Albania)\\b"
dz	<- "\\b(?i)(Algeria)\\b"
as	<- "\\b(?i)(American Samoa)\\b"
ad	<- "\\b(?i)(Andorra)\\b"
ao	<- "\\b(?i)(Angola)\\b"
ai	<- "\\b(?i)(Anguilla)\\b"
aq	<- "\\b(?i)(Antarctica|Antartica)\\b"
ag	<- "\\b(?i)(Antigua|Barbuda)\\b"
ar	<- "\\b(?i)(Argentina|Buenos Aires|La plata)\\b"
am	<- "\\b(?i)(Armenia)\\b"
aw	<- "\\b(?i)(Aruba)\\b"
au	<- "\\b(?i)(Australia|Melbourne(?!, )FL|Sydney(?!, Nova)|Wollongong|Brisbane|Canberra)\\b"
at  <- "\\b(?i)(Austria|Vienna(?!, V))\\b"
az	<- "\\b(?i)(Azerbaijan)\\b"
bh	<- "\\b(?i)(Bahrain)\\b"
bs	<- "\\b(?i)(Bahamas)\\b"
bd	<- "\\b(?i)(Bangladesh|Dhaka)\\b"
bb	<- "\\b(?i)(Barbados)\\b"
by	<- "\\b(?i)(Belarus)\\b"
be  <- "\\b(?i)(Belgium|Brussels)\\b"
bz	<- "\\b(?i)(Belize)\\b"
bj	<- "\\b(?i)(Benin(?! City))\\b"
bm	<- "\\b(?i)(Bermuda)\\b"
bt	<- "\\b(?i)(Bhutan)\\b"
bo	<- "\\b(?i)(Bolivia)\\b"
bq	<- "\\b(?i)(Bonaire)\\b"
ba	<- "\\b(?i)(Bosnia|Herzegovina)\\b"
bw	<- "\\b(?i)(Botswana)\\b"
bv	<- "\\b(?i)(Bouvet Island)\\b"
br  <- "\\b(?i)(Brazil|Brasil|Aracaju|Sao Paulo|San Paulo|Rio de Janeiro|Brasilia|Braslia|Fortaleza|Araraquara|Barueri|Belo Horizonte|BR,|Brazilian|Campina Grande|Florianpolis|Fortaleza|Goinia-GO|Goinia|Joo Pessoa|Porto Alegre|Rio de |Rio Grande|So Jos dos|So Paulo)\\b"
bn	<- "\\b(?i)(Brunei Darussalam|Brunei)\\b"
bg  <- "\\b(?i)(Bulgaria)\\b"
bf	<- "\\b(?i)(Burkina Faso)\\b"
bi	<- "\\b(?i)(Burundi)\\b"
kh	<- "\\b(?i)(Cambodia)\\b"
cm	<- "\\b(?i)(Cameroon)\\b"
ca  <- "\\b(?i)(Canada|Ontario(?!, CA)|Quebec|nova scotia|new brunswick(?!, N)|manitoba|british columbia|, BC|prince edward island|saskatchewan|alberta|newfoundland|toronto|Vancouver(?!\\, W)|, ON|Winnipeg|, B.C.|,ON|Victoria|Burnaby|Fredericton|Montreal|Qubec)\\b"
cv	<- "\\b(?i)(Cape Verde)\\b"
ky	<- "\\b(?i)(Cayman Islands)\\b"
cf	<- "\\b(?i)(Central African Republic)\\b"
td	<- "\\b(?i)(Chad)\\b"
cl	<- "\\b(?i)(Chile)\\b"
cn  <- "\\b(?i)(China|Beijing|GuangZhou|Shanghai|Tianjin|Shenzhen|Chengdu|Dongguan|Chongqing|Guangdong|ZhuHaiChina|Zhu haiChina|Zhengzhou|ZheJiang|HangZhou|zh-cn|(?<![Zurich|Lausanne]), CH|, CHN|Xiamen|Wuhan|beijingchina|Changsha|Hunan|ChengduChina|chinabeijing|ChinaShenzhen|^guang|^Hangz|Harbin Institute|Hefei|Hunan|^jiang|^Nanjing|^ShenZhen)\\b"
cx	<- "\\b(?i)(Christmas Island)\\b"
cc	<- "\\b(?i)(Cocos Islands)\\b"
co	<- "\\b(?i)(Colombia|Bogataa|Medellin)\\b"
km	<- "\\b(?i)(Comoros)\\b"
cg	<- "\\b(?i)((?<!Democratic Republic of the )Congo)\\b"
cd	<- "\\b(?i)(Democratic Republic of the Congo|Brazzaville)\\b"
ck	<- "\\b(?i)(Cook Islands)\\b"
cr	<- "\\b(?i)(Costa Rica)\\b"
ci	<- "\\b(?i)(Cte dIvoire|Cote dIvoire|Cote d'Ivoire)\\b"
hr	<- "\\b(?i)(Croatia|Zagreb)\\b"
cu	<- "\\b(?i)(Cuba)\\b"
cw	<- "\\b(?i)(Curaao)\\b"
cy	<- "\\b(?i)(Cyprus)\\b"
cz	<- "\\b(?i)(Czech Republic|Czech|Czechia|Prague|Brno|CzechRepublic)\\b"
dk	<- "\\b(?i)(Denmark|Copenhagen)\\b"
dj	<- "\\b(?i)(Djibouti)\\b"
dm	<- "\\b(?i)(Dominica)\\b"
do	<- "\\b(?i)(Dominican Republic|, DR|dominican|Dominican)\\b"
ec	<- "\\b(?i)(Ecuador)\\b"
eg	<- "\\b(?i)(Egypt)\\b"
sv	<- "\\b(?i)(El Salvador)\\b"
gq	<- "\\b(?i)(Equatorial Guinea)\\b"
er	<- "\\b(?i)(Eritrea)\\b"
ee	<- "\\b(?i)(Estonia)\\b"
et	<- "\\b(?i)(Ethiopia)\\b"
fk	<- "\\b(?i)(Falkland Islands|Malvinas)\\b"
fo	<- "\\b(?i)(Faroe Islands|Faroe)\\b"
fj	<- "\\b(?i)(Fiji)\\b"
fi	<- "\\b(?i)(Finland|Helsinki)\\b"
fr	<- "\\b(?i)(France|Paris|, FR|Toulouse|French Riviera|French Alps|Lille|Lyon|Sophia Antipolis)\\b"
gf	<- "\\b(?i)(French Guiana)\\b"
pf	<- "\\b(?i)(French Polynesia)\\b"
tf	<- "\\b(?i)(French Southern Territories)\\b"
ga	<- "\\b(?i)(Gabon)\\b"
gm	<- "\\b(?i)(Gambia)\\b"
# we decided to code instances of "georgia" as us since there was no easy way to demarcate and most seemed like they from the us 
ge	<- "\\b(?i)(Tbilisi|Batumi|Kutaisi)\\b" #"\\b(?i)((?<!, )Georgia(?!, U)|Tbilisi)\\b"
de <- "\\b(?i)(Germany|German|Deutschland|Berlin|Bavaria|Bayern|Munich|Dresden|Frankfurt|Hamburg)\\b"
gh	<- "\\b(?i)(Ghana)\\b"
gi	<- "\\b(?i)(Gibraltar)\\b"
gr	<- "\\b(?i)(Greece)\\b"
gl	<- "\\b(?i)(Greenland)\\b"
gd	<- "\\b(?i)(Grenada)\\b"
gp	<- "\\b(?i)(Guadeloupe)\\b"
gu	<- "\\b(?i)(Guam)\\b"
gt	<- "\\b(?i)(Guatemala)\\b"
gg	<- "\\b(?i)(Guernsey)\\b"
gn	<- "\\b(?i)(Guinea)\\b"
gw	<- "\\b(?i)(Guinea-Bissau)\\b"
gy	<- "\\b(?i)(Guyana)\\b"
ht	<- "\\b(?i)(Haiti)\\b"
hm	<- "\\b(?i)(Heard Island|McDonald Islands)\\b"
va	<- "\\b(?i)(Holy See|Vatican)\\b"
hn	<- "\\b(?i)(Honduras)\\b"
hk	<- "\\b(?i)(Hong Kong)\\b"
hu	<- "\\b(?i)(Hungary|Budapest)\\b"
is	<- "\\b(?i)(Iceland)\\b"
id	<- "\\b(?i)(Indonesia|Yogyakarta|Bandung|Jakarta)\\b"
ir	<- "\\b(?i)(Iran)\\b"
iq	<- "\\b(?i)(Iraq)\\b"
ie	<- "\\b(?i)(Ireland|Dublin(?!, [OC]))\\b"
im	<- "\\b(?i)(Isle of Man)\\b"
il	<- "\\b(?i)(Israel)\\b"
it	<- "\\b(?i)(Italy|Rome(?!, GA)|Milan|Milano|Turin|Trieste|Bologna)\\b"
jm	<- "\\b(?i)(Jamaica(?! Plain))\\b"
jp  <- "\\b(?i)(Japan|Tokyo|Kyoto|Osaka|Yokohama|ja|ja_jp|ja-jp|Tsukuba)\\b"
jo	<- "\\b(?i)(Jordan(?!, U))\\b"
kz	<- "\\b(?i)(Kazakhstan)\\b"
ke	<- "\\b(?i)(Kenya|Nairobi)\\b"
ki	<- "\\b(?i)(Kiribati)\\b"
kp	<- "\\b(?i)(North Korea|Pyongyang)\\b"
kr	<- "\\b(?i)(South Korea|Seoul|Seuol|Daejeon|pangyo)\\b"
xk  <- "\\b(?i)(Kosovo|Kosova|^Prishtina|^Prishtine|^Pristina)\\b"
kw	<- "\\b(?i)(Kuwait)\\b"
kg	<- "\\b(?i)(Kyrgyzstan)\\b"
la	<- "\\b(?i)(Lao Peoples Democratic Republic|Laos)\\b"
lv <- "\\b(?i)(Latvia)\\b"
lb	<- "\\b(?i)(Lebanon(?! , NH))\\b"
ls	<- "\\b(?i)(Lesotho)\\b"
lr	<- "\\b(?i)(Liberia)\\b"
ly	<- "\\b(?i)(Libya)\\b"
li	<- "\\b(?i)(Liechtenstein)\\b"
lu <- "\\b(?i)(Luxembourg)\\b"
lt <- "\\b(?i)(Lithuania)\\b"
mo	<- "\\b(?i)(Macao)\\b"
mk	<- "\\b(?i)(Macedonia)\\b"
mg	<- "\\b(?i)(Madagascar)\\b"
mw	<- "\\b(?i)(Malawi)\\b"
my	<- "\\b(?i)(Malaysia)\\b"
mv	<- "\\b(?i)(Maldives)\\b"
ml	<- "\\b(?i)(Mali)\\b"
mt <- "\\b(?i)(Malta)\\b"
mh	<- "\\b(?i)(Marshall Islands)\\b"
mq	<- "\\b(?i)(Martinique)\\b"
mr	<- "\\b(?i)(Mauritania)\\b"
mu	<- "\\b(?i)(Mauritius)\\b"
yt	<- "\\b(?i)(Mayotte)\\b"
mx	<- "\\b(?i)((?<!New )Mexico|(?<!New )Mxico|Zacatecas|Mex.|, MX|Villahermosa|Colima)\\b"
fm	<- "\\b(?i)(Micronesia)\\b"
md <- "\\b(?i)(Moldova)\\b"
mc <- "\\b(?i)(Monaco)\\b"
mn	<- "\\b(?i)(Mongolia)\\b"
me	<- "\\b(?i)(Montenegro)\\b"
ms	<- "\\b(?i)(Montserrat)\\b"
ma	<- "\\b(?i)(Morocco)\\b"
mz	<- "\\b(?i)(Mozambique)\\b"
mm	<- "\\b(?i)(Myanmar)\\b"
na	<- "\\b(?i)(Namibia)\\b"
nr	<- "\\b(?i)(Nauru)\\b"
np	<- "\\b(?i)(Nepal)\\b"
nl <- "\\b(?i)(Netherlands|Amsterdam|Holland(?!, [MO])|Zwolle|Willemstad|Curacao|Curaao|Utrecht|Breda|Eindhoven|Groningen)\\b"
nc	<- "\\b(?i)(New Caledonia)\\b"
nz	<- "\\b(?i)(New Zealand|Auckland|NZ|NZL)\\b"
ni	<- "\\b(?i)(Nicaragua)\\b"
ne	<- "\\b(?i)(Niger)\\b"
ng	<- "\\b(?i)(Nigeria|Lagos)\\b"
nu	<- "\\b(?i)(Niue)\\b"
nf	<- "\\b(?i)(Norfolk Island)\\b"
mp	<- "\\b(?i)(Northern Mariana Islands)\\b"
no	<- "\\b(?i)(Norway)\\b"
om	<- "\\b(?i)(Oman)\\b"
pk	<- "\\b(?i)(Pakistan)\\b"
pw	<- "\\b(?i)(Palau)\\b"
ps	<- "\\b(?i)(Palestine|Gaza)\\b"
pa	<- "\\b(?i)(Panama(?! City))\\b"
pg	<- "\\b(?i)(Papua New Guinea)\\b"
py	<- "\\b(?i)(Paraguay)\\b"
pe	<- "\\b(?i)(Peru|Arequipa)\\b"
ph	<- "\\b(?i)(Philippines)\\b"
pn	<- "\\b(?i)(Pitcairn)\\b"
pl <- "\\b(?i)(Poland|Warsaw|Wrocaw)\\b"
pt <- "\\b(?i)(Portugal|Lisboa|Lisbon|Azores|Porto(?! Alegre)|Faro)\\b"
pr	<- "\\b(?i)(Puerto Rico)\\b"
qa	<- "\\b(?i)(Qatar)\\b"
re	<- "\\b(?i)(Runion)\\b"
ro <- "\\b(?i)(Romania|Bucharest)\\b"
ru <- "\\b(?i)(Russia|Russian|Moscow(?!, I)|Yekaterinburg|Petersburg(?!, F)|St-Petersburg(?!, F)|RU, )\\b"
rw	<- "\\b(?i)(Rwanda)\\b"
bl	<- "\\b(?i)(Saint Barthlemy)\\b"
sh	<- "\\b(?i)(Saint Helena)\\b"
kn	<- "\\b(?i)(Saint Kitts|Nevis)\\b"
lc	<- "\\b(?i)(Saint Lucia)\\b"
mf	<- "\\b(?i)(Saint Martin)\\b"
pm	<- "\\b(?i)(Saint Pierre|Miquelon)\\b"
vc	<- "\\b(?i)(Saint Vincent|Grenadines)\\b"
ws	<- "\\b(?i)(Samoa)\\b"
sm	<- "\\b(?i)(San Marino)\\b"
st	<- "\\b(?i)(Sao Tome|Principe)\\b"
sa	<- "\\b(?i)(Saudi Arabia)\\b"
sn <- "\\b(?i)(Senegal)\\b"
rs	<- "\\b(?i)(Serbia)\\b"
sc	<- "\\b(?i)(Seychelles)\\b"
sl	<- "\\b(?i)(Sierra Leone)\\b"
sg	<- "\\b(?i)(Singapore)\\b"
sx	<- "\\b(?i)(Sint Maarten)\\b"
sk <- "\\b(?i)(Slovak Republic|Slovakia)\\b"
si <- "\\b(?i)(Slovenia)\\b"
sb	<- "\\b(?i)(Solomon Islands)\\b"
so	<- "\\b(?i)(Somalia)\\b"
za	<- "\\b(?i)(South Africa|Cape Town|Johannesburg)\\b"
gs	<- "\\b(?i)(South Georgia|South Sandwich Islands|South Sandwich|Sandwich Islands)\\b"
ss	<- "\\b(?i)(South Sudan)\\b"
es <- "\\b(?i)(Spain|Madrid|Barcelona|Granada|Sevilla|Mallorca|Ibiza|Bilbao|Espana|Espaa|Zaragoza|Catalunya|Basque|Valencia(?!, [C|Ven])|Catalonia|Galicia)\\b"
lk	<- "\\b(?i)(Sri Lanka|Srilanka)\\b"
sd	<- "\\b(?i)(Sudan)\\b"
sr	<- "\\b(?i)(Suriname)\\b"
sj	<- "\\b(?i)(Svalbard|Jan Mayen)\\b"
sz	<- "\\b(?i)(Swaziland)\\b"
se <- "\\b(?i)(Sweden|Stockholm)\\b"
ch	<- "\\b(?i)(Switzerland|Zurich|Zrich|Lausanne)\\b"
sy	<- "\\b(?i)(Syrian Arab Republic|syria)\\b"
tw	<- "\\b(?i)(Taiwan|^Taipei)\\b"
tj	<- "\\b(?i)(Tajikistan)\\b"
tz	<- "\\b(?i)(Tanzania)\\b"
th	<- "\\b(?i)(Thailand)\\b"
tl	<- "\\b(?i)(Timor-Leste)\\b"
tg	<- "\\b(?i)(Togo)\\b"
tk	<- "\\b(?i)(Tokelau)\\b"
to	<- "\\b(?i)(Tonga)\\b"
tt	<- "\\b(?i)(Trinidad|Tobago)\\b"
tn	<- "\\b(?i)(Tunisia)\\b"
tr	<- "\\b(?i)(Turkey|Istanbul|Trkiye|Estambul)\\b"
tm	<- "\\b(?i)(Turkmenistan)\\b"
tc	<- "\\b(?i)(Turks|Caicos Islands)\\b"
tv	<- "\\b(?i)(Tuvalu)\\b"
ug	<- "\\b(?i)(Uganda)\\b"
ua	<- "\\b(?i)(Ukraine|Ukrane|Kyiv|Kiev|Zaporozhye|Dnepr|Dnipro)\\b"
ae	<- "\\b(?i)(United Arab|Emirates|UAE|Abu Dhabi)\\b"
gb <- "\\b(?i)(UK|UK,|United Kingdom|united-kingdom|London|(?<!New )England|Wales|U.K.|Aberdeen|Cardiff(?!, CA)|Scotland|Edinburgh|Glasglow|Brighton(?!, [M|C]))\\b"
us	<- "\\b(?i)(?:Ala(?:(?:bam|sk)a)|American Samoa|Arizona|Arkansas|California|Colorado|Connecticut|Delaware|District of Columbia|Florida|Georgia|Guam|Hawaii|Idaho|Illinois|Indiana|Iowa|Kansas|Kentucky|Louisiana|Maine|Maryland|Massachusetts|Michigan|Minnesota|Miss(?:(?:issipp|our)i)|Montana|Nebraska|Nevada|New (?:Hampshire|Jersey|Mexico|York)|North (?:(?:Carolin|Dakot)a)|Ohio|Oklahoma|Oregon|Pennsylvania|Puerto Rico|Rhode Island|South (?:(?:Carolin|Dakot)a)|Tennessee|Texas|Utah|Vermont|Virgin(?:ia| Island(s?))|Washington|West Virginia|Wisconsin|Wyoming|United States|A[KLRSZ]|C[AOT]|D[CE]|FL|G[AU]|HI|I[ADLN]|K[SY]|LA|M[ADEINOST]|N[CDEHJMVY]|O[HKR]|P[AR]|RI|S[CD]|T[NX]|UT|V[AIT]|W[AIVY]|NYC|DC|D.C.|US|USA|Seattle|San Francisco|Chicago|Philly|Philadelphia|Los Angeles|Portland|Houston|Phoenix|San Antonio|San Diego|Oakland|Boston|Bay Area|Berkeley|Stanford|Harvard|Dallas|Silicon|San Jose|USC|Twin Cities|NYU|Yale|, Mass| Mass|newyork|new_york|U.S.|Atlanta|Austin|Albany|Manhattan|Brooklyn|Bronx|Cal|Cape Cod|Cleveland|Denver|Detroit|East Bay|Pittsburgh|Salt Lake|St. Louis|Honolulu|Miami|Minneapolis|Palo Alto|^san fran|socal)\\b"
uy	<- "\\b(?i)(Uruguay)\\b"
uz	<- "\\b(?i)(Uzbekistan)\\b"
vu	<- "\\b(?i)(Vanuatu)\\b"
ve	<- "\\b(?i)(Venezuela)\\b"
vn	<- "\\b(?i)(Viet Nam|Vietnam)\\b"
vg	<- "\\b(?i)(Virgin Islands, British|British Virgin Islands)\\b"
vi	<- "\\b(?i)(Virgin Islands, U.S.|Virgin Islands, US| US Virgin Islands)\\b"
wf	<- "\\b(?i)(Wallis and Futuna)\\b"
eh	<- "\\b(?i)(Western Sahara)\\b"
ye	<- "\\b(?i)(Yemen)\\b"
zm	<- "\\b(?i)(Zambia)\\b"
zw	<- "\\b(?i)(Zimbabwe)\\b"

text_data <- text_data %>%
  mutate(department = removePunctuation(department)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, 
                                   pattern = af), yes = "af", no = "")) %>% 
                                   #paste("af", detected_country, sep=","), no = "")) 
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ax), 
                                   paste("ax", detected_country, sep=","), 
                                   no = detected_country)) %>% 
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = al), 
                                   paste("al", detected_country, sep=","), 
                                   no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = dz), 
                                   paste("dz", detected_country, sep=","), 
                                   no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = as), 
                                   paste("as", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ad), 
                                   paste("ad", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ao), 
                                   paste("ao", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ai), 
                                   paste("ai", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = aq), 
                                   paste("aq", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ag), 
                                   paste("ag", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ar), 
                                   paste("ar", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = am), 
                                   paste("am", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = aw), 
                                   paste("aw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = au), 
                                   paste("au", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = at), 
                                   paste("at", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = be), 
                                   paste("be", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bg), 
                                   paste("bg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = br), 
                                   paste("br", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = az), 
                                   paste("az", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bh), 
                                   paste("bh", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bs), 
                                   paste("bs", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bd), 
                                   paste("bd", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bb), 
                                   paste("bb", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = by), 
                                   paste("by", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bz), 
                                   paste("bz", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bj), 
                                   paste("bj", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bm), 
                                   paste("bm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bt), 
                                   paste("bt", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bo), 
                                   paste("bo", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bq), 
                                   paste("bq", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ba), 
                                   paste("ba", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bw), 
                                   paste("bw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bv), 
                                   paste("bv", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bn), 
                                   paste("bn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bf), 
                                   paste("bf", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bi), 
                                   paste("bi", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cy), 
                                   paste("cy", detected_country, sep=","), no = detected_country)) %>% 
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cz), 
                                   paste("cz", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cn), 
                                   paste("cn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = kh), 
                                   paste("kh", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cm), 
                                   paste("cm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ca), 
                                   paste("ca", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cv), 
                                   paste("cv", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ky), 
                                   paste("ky", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cf), 
                                   paste("cf", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = td), 
                                   paste("td", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cl), 
                                   paste("cl", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cx), 
                                   paste("cx", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cc), 
                                   paste("cc", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = co), 
                                   paste("co", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = km), 
                                   paste("km", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ck), 
                                   paste("ck", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cr), 
                                   paste("cr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ci), 
                                   paste("ci", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = hr), 
                                   paste("hr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cu), 
                                   paste("cu", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = cw), 
                                   paste("cw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = dk), 
                                   paste("dk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ee), 
                                   paste("ee", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = fi), 
                                   paste("fi", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = fr), 
                                   paste("fr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = dj), 
                                   paste("dj", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = dm), 
                                   paste("dm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = do), 
                                   paste("do", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ec), 
                                   paste("ec", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = eg), 
                                   paste("eg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sv), 
                                   paste("sv", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gq), 
                                   paste("gq", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = er), 
                                   paste("er", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = et), 
                                   paste("et", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = fk), 
                                   paste("fk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = fo), 
                                   paste("fo", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = fj), 
                                   paste("fj", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = de), 
                                   paste("de", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gr), 
                                   paste("gr", detected_country, sep=","), no = detected_country)) %>% 
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gf), 
                                   paste("gf", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pf), 
                                   paste("pf", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tf), 
                                   paste("tf", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ga), 
                                   paste("ga", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gm), 
                                   paste("gm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ge), 
                                   paste("ge", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gh), 
                                   paste("gh", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gi), 
                                   paste("gi", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gl), 
                                   paste("gl", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gd), 
                                   paste("gd", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gp), 
                                   paste("gp", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gu), 
                                   paste("gu", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gt), 
                                   paste("gt", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gg), 
                                   paste("gg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gn), 
                                   paste("gn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gw), 
                                   paste("gw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gy), 
                                   paste("gy", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = hu), 
                                   paste("hu", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ie), 
                                   paste("ie", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = it), 
                                   paste("it", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ht), 
                                   paste("ht", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = hm), 
                                   paste("hm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = va), 
                                   paste("va", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = hn), 
                                   paste("hn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = hk), 
                                   paste("hk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = is), 
                                   paste("is", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = id), 
                                   paste("id", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ir), 
                                   paste("ir", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = iq), 
                                   paste("iq", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = im), 
                                   paste("im", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = il), 
                                   paste("il", detected_country, sep=","), no = detected_country)) %>%
  # i did india down here so that i didn't have to name a variable "in", which created problems 
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = "\\b(?i)(India|Mumbai|Delhi|Kolkata|Chennai|Bangalore|Hyderabad|Ahmedabad|Pune|VIT |NIT |Amritsar|(?<!Lahore[ ,])Punjab(?! Pakistan)|Bhubaneswar|Gandhinagar|Gujarat|^IIT |Indian|Jaipur|jammu|kashmir|^MNNIT)\\b"), 
                                      paste("in", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = jp), 
                                   paste("jp", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = lv), 
                                   paste("lv", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = lu), 
                                   paste("lu", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = lt), 
                                   paste("lt", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = jm), 
                                   paste("jm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = jo), 
                                   paste("jo", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = kz), 
                                   paste("kz", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ke), 
                                   paste("ke", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ki), 
                                   paste("ki", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = kp), 
                                   paste("kp", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = kr), 
                                   paste("kr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = kw), 
                                   paste("kw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = kg), 
                                   paste("kg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = xk), 
                                   paste("xk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = la), 
                                   paste("la", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = lb), 
                                   paste("lb", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ls), 
                                   paste("ls", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = lr), 
                                   paste("lr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ly), 
                                   paste("ly", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = li), 
                                   paste("li", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mt), 
                                   paste("mt", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mo), 
                                   paste("mo", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mk), 
                                   paste("mk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mg), 
                                   paste("mg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mw), 
                                   paste("mw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = my), 
                                   paste("my", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mv), 
                                   paste("mv", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ml), 
                                   paste("ml", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mh), 
                                   paste("mh", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mq), 
                                   paste("mq", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mr), 
                                   paste("mr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mu), 
                                   paste("mu", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mx), 
                                   paste("mx", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = yt), 
                                   paste("yt", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = fm), 
                                   paste("fm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = md), 
                                   paste("md", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mc), 
                                   paste("mc", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mn), 
                                   paste("mn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = me), 
                                   paste("me", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ms), 
                                   paste("ms", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ma), 
                                   paste("ma", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mz), 
                                   paste("mz", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mm), 
                                   paste("mm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = na), 
                                   paste("na", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = nr), 
                                   paste("nr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = np), 
                                   paste("np", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = nl), 
                                   paste("nl", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = nc), 
                                   paste("nc", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = nz), 
                                   paste("nz", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ni), 
                                   paste("ni", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ne), 
                                   paste("ne", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ng), 
                                   paste("ng", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = nu), 
                                   paste("nu", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = nf), 
                                   paste("nf", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mp), 
                                   paste("mp", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = no), 
                                   paste("no", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = om), 
                                   paste("om", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pk), 
                                   paste("pk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pw), 
                                   paste("pw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ps), 
                                   paste("ps", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pa), 
                                   paste("pa", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pg), 
                                   paste("pg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = py), 
                                   paste("py", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pe), 
                                   paste("pe", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ph), 
                                   paste("ph", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pn), 
                                   paste("pn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pl), 
                                   paste("pl", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pt), 
                                   paste("pt", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pr), 
                                   paste("pr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = qa), 
                                   paste("qa", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = re), 
                                   paste("re", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ro), 
                                   paste("ro", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ru), 
                                   paste("ru", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = rw), 
                                   paste("rw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = bl), 
                                   paste("bl", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sh), 
                                   paste("sh", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = kn), 
                                   paste("kn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = lc), 
                                   paste("lc", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = mf), 
                                   paste("mf", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = pm), 
                                   paste("pm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = vc), 
                                   paste("vc", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ws), 
                                   paste("ws", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sm), 
                                   paste("sm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = st), 
                                   paste("st", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sa), 
                                   paste("sa", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sn), 
                                   paste("sn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = rs), 
                                   paste("rs", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sc), 
                                   paste("sc", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sl), 
                                   paste("sl", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sg), 
                                   paste("sg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sx), 
                                   paste("sx", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sk), 
                                   paste("sk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = si), 
                                   paste("si", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sb), 
                                   paste("sb", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = so), 
                                   paste("so", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = za), 
                                   paste("za", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gs), 
                                   paste("gs", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ss), 
                                   paste("ss", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = es), 
                                   paste("es", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = lk), 
                                   paste("lk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sd), 
                                   paste("sd", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sr), 
                                   paste("sr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sj), 
                                   paste("sj", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sz), 
                                   paste("sz", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = se), 
                                   paste("se", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ch), 
                                   paste("ch", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = sy), 
                                   paste("sy", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tw), 
                                   paste("tw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tj), 
                                   paste("tj", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tz), 
                                   paste("tz", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = th), 
                                   paste("th", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tl), 
                                   paste("tl", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tg), 
                                   paste("tg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tk), 
                                   paste("tk", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = to), 
                                   paste("to", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tt), 
                                   paste("tt", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tn), 
                                   paste("tn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tr), 
                                   paste("tr", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tm), 
                                   paste("tm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tc), 
                                   paste("tc", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = tv), 
                                   paste("tv", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ug), 
                                   paste("ug", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ua), 
                                   paste("ua", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ae), 
                                   paste("ae", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = gb), 
                                   paste("gb", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = uy), 
                                   paste("uy", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = uz), 
                                   paste("uz", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = vu), 
                                   paste("vu", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ve), 
                                   paste("ve", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = vn), 
                                   paste("vn", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = vg), 
                                   paste("vg", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = vi), 
                                   paste("vi", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = wf), 
                                   paste("wf", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = eh), 
                                   paste("eh", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = ye), 
                                   paste("ye", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = zm), 
                                   paste("zm", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = zw), 
                                   paste("zw", detected_country, sep=","), no = detected_country)) %>%
  mutate(detected_country = ifelse(test = str_detect(string = department, pattern = us), 
                                   paste("us", detected_country, sep=","), no = detected_country)) %>% 
  dplyr::rename(country_code = detected_country)

text_data <- text_data %>% 
  separate_rows(country_code, sep = "," , convert = FALSE) %>% 
  mutate(country_code = trimws(country_code)) %>% 
  mutate(country_code = na_if(country_code, ""))

# removing the commas at the end of the strings from the column just created
text_data$country_code <- gsub("\\,$", "", text_data$country_code)
text_data$country_code <- trimws(text_data$country_code)

text_data$country <- countrycode(text_data$country_code, 
                                     origin = "iso2c", destination = "country.name")

# tokenizing the abstract data into words 
abstract_data_Cc <- text_data %>% 
  unnest_tokens(word, abstract) %>% 
  anti_join(stop_words)
# adding custom set of stopwords 
my_stopwords <- tibble(word = c(as.character(1:14), 
                                "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", 
                                "rights", "reserved", "copyright", "elsevier"))
abstract_data_Cc <- abstract_data_Cc %>% anti_join(my_stopwords); rm(my_stopwords)
# most frequent word count in abstracts 
abstract_data_Cc %>%
  count(word, sort = TRUE)

```

```{r figure 6a, fig.width=9.5, echo=FALSE}
# standardizing all countries  
pop_terms_bycountry <- abstract_data_Cc %>% 
  separate(country, into = c("country", "void"), 
           sep = ",", extra = "drop", fill = "right") %>% 
  select(word, country, country_code, year, pubmed_id) %>%    
  #mutate(country = stri_trim_both(country_code)) %>% 
  #mutate(country = tolower(country_code)) %>%
  mutate(diversity = ifelse(test = str_detect(string = word, 
         pattern = "diversity"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
         pattern = all_pop_terms), 
         yes = "all population terms", no = word)) %>%  
  mutate(all_pop_terms = ifelse(test = str_detect(string = term, 
         pattern = "\\b(all population terms)\\b"), yes = 1, no = 0)) 

bycountry_graph <- pop_terms_bycountry %>%   
    drop_na(country) %>%
  group_by(year) %>% 
  count(term, country, sort = TRUE) %>% 
  ungroup() %>% 
  filter(term == "all population terms" ) %>% 
  filter(  country == "United States" | country == "China" | country == "United Kingdom" | 
           country == "Germany" | country == "Australia" 
  #         | country == "France" | country == "Canada" | country == "Netherlands" 
  #         | country == "Australia" | country == "Spain"   
           ) %>%  
  ggplot() + geom_line(aes(y = n, x = year, colour = country), stat="identity") +
  labs(title = "Figure 6A. Growth in Diversity-Related Terminology by Country (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

```{r figure 6b, fig.width=9.5, echo=FALSE}
# total articles each year 
bycountry_prc_counts <- text_data %>% 
  drop_na(country) %>% 
  group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n)  

#abstract_data_Cc %>% 
#  filter(year == 1990) %>% 
#  mutate(department = removePunctuation(department)) %>% 
#  select(author, year, department, country)   

bycountry_prc_counts <- pop_terms_bycountry %>% 
  filter(all_pop_terms == 1) %>% 
  group_by(country, year) %>% 
  summarise(cnt_pop_terms = n_distinct(pubmed_id)) %>% 
  right_join(bycountry_prc_counts, by = "year") %>% 
  mutate(prc_pop_terms = round(cnt_pop_terms / total * 100, digits = 2))
# graphing 
bycountry_graph <- bycountry_prc_counts %>% 
  drop_na(country) %>% 
  filter(  country == "United States" | country == "China" | country == "United Kingdom" | 
           country == "Germany" | country == "Australia" 
  #         | country == "France" | country == "Canada" | country == "Netherlands" 
  #         | country == "Australia" | country == "Spain"   
           ) %>% 
  ggplot() + 
  geom_line(aes(y = prc_pop_terms, x = year, colour = country), stat="identity") +
  labs(title = "Figure 6B. Growth in Diversity-Related Terminology by Country (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()

```

When we look at growth in all population terms (see the full list from above) as a percentage by country, we see very similar patterns as we did with the raw counts. The only difference is that population terms show up in a larger percentage of abstracts deriving from the US and England in the 1990s while the growth of population terminology only increases in the Netherlands after 2006. 

### Examining Diversity Across Academic Disciplines 

Lastly, we wanted to look more into the growth of diversity-related terms by scientific subject matter. This analysis breaks down the number of words occurring in abstracts over time, which is then broken down by MEDLINE's and Web of Science's subject categories. I have opted to only include 12 of the 150 categories that could have been graphed here.  

```{r figure 7a, message = FALSE, warning = FALSE, fig.width=9.5, echo=FALSE}
# WARNING: You will need at least 32GB of RAM to run this chunk so let's make as much space as we can.
rm(pubs_by_year, us_pop_terms, us_pop_counts, us_pop_prc_counts, 
   all_pop_prc_counts, all_pop_terms_graph, all_pop_terms_trends, 
   gen_pop_prc_counts, pop_terms_bycats, bycats_prc_counts,
   bycountry_graph, bycountry_prc_counts, pop_terms_bycountry)
# standardizing all subjects  
pop_terms_bysubject <- abstract_data %>% 
  separate(subject, into = paste("subject", 1:15, sep = "_"), sep = ";") %>%
  gather(value, subject, subject_1:subject_15, na.rm = TRUE) %>% select(-value) %>% 
  separate(subject, into = c("subject", "void"), sep = "[(]") %>% select(-void) %>% 
  mutate(subject = stri_trim_both(subject)) %>% 
  mutate(subject = tolower(subject)) %>% 
  mutate(diversity = ifelse(test = str_detect(string = word, 
         pattern = "diversity"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = word, 
         pattern = all_pop_terms), 
         yes = "all population terms", no = word)) %>%  
  mutate(all_pop_terms = ifelse(test = str_detect(string = term, 
         pattern = "\\b(all population terms)\\b"), yes = 1, no = 0))
# graphing 
bysubject_graph <- pop_terms_bysubject %>%   
  select(subject, term, year) %>%
  group_by(year) %>% 
  count(term, subject, sort = TRUE) %>% ungroup() %>%     
  filter(term == "all population terms") %>% 
  filter(subject == "genetics & heredity" | subject == "biochemistry & molecular biology" | 
         subject == "microbiology" | subject == "infectious diseases" | subject == "immunology" | 
         subject == "pharmacology & pharmacy" | subject == "behavioral sciences" |
         subject == "health care sciences & services" | subject == "neurosciences & neurology" |
         subject == "psychology" | subject == "sociology" | 
         subject == "oncology" | subject == "business & economics" ) %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = subject), stat="identity") +
  labs(title = "Figure 7A. Growth in Diversity-Related Terminology by Select Subject (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

Overall, we see that the rise of diversity in genetics & heredity, biochemistry & molecular biology, microbiology, immunology, and infectious disease research. While there is also an increased use of diversity-related terminology in the social sciences (like psychology, sociology and behavioral sciences), we generally see the magnitude of this rise is more muted. In other cases, like business & economics and neuroscience & neurology, we see almost no increase at all.

```{r figure 7b, message = FALSE, warning = FALSE, fig.width=9.5, echo=FALSE}
# total articles each year 
bysubject_prc_counts <- text_data %>% group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n) 
# articles with term mentioned each year 
bysubject_prc_counts <- pop_terms_bysubject %>% filter(diversity == 1) %>% 
  group_by(subject, year) %>% summarise(cnt_diversity = n_distinct(id)) %>% 
  right_join(bysubject_prc_counts, by = "year") %>% 
  mutate(prc_diversity = round(cnt_diversity / total * 100, digits = 2)) 
# wrangling all cats into a df 
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "genetics & heredity") %>% 
  mutate(genetics_heredity = prc_diversity) 
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "biochemistry & molecular biology") %>% 
  mutate(biochem_molbio = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "microbiology") %>% 
  mutate(microbio = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "infectious diseases") %>% 
  mutate(infectious = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "immunology") %>% 
  mutate(immunology = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "pharmacology & pharmacy") %>% 
  mutate(pharmacology = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "behavioral sciences") %>% 
  mutate(behavioral = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "health care sciences & services") %>% 
  mutate(healthcare = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "neurosciences & neurology") %>% 
  mutate(neurology = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "psychology") %>% 
  mutate(psychology = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "sociology") %>% 
  mutate(sociology = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "oncology") %>% 
  mutate(oncology = prc_diversity) %>% right_join(bysubject_graph, by = "year")
bysubject_graph <- bysubject_prc_counts %>% filter(subject == "business & economics") %>% 
  mutate(business = prc_diversity) %>% right_join(bysubject_graph, by = "year")
# graphing top-10ish countries over time 
bysubject_graph <- bysubject_graph %>% ggplot() + 
  geom_line(aes(y = genetics_heredity, x = year, colour = "genetics & heredity"), stat="identity") +
  geom_line(aes(y = biochem_molbio, x = year, 
                colour = "biochemistry & molecular biology"), stat="identity") +
  geom_line(aes(y = microbio, x = year, 
                colour = "microbiology"), stat="identity") +
  geom_line(aes(y = infectious, x = year, 
                colour = "infectious diseases"), stat="identity") +
  geom_line(aes(y = immunology, x = year, 
                colour = "immunology"), stat="identity") +
  geom_line(aes(y = pharmacology, x = year, 
                colour = "pharmacology & pharmacy"), stat="identity") +
  geom_line(aes(y = behavioral, x = year, 
                colour = "behavioral sciences"), stat="identity") +
  geom_line(aes(y = healthcare, x = year, 
                colour = "health care sciences & services"), stat="identity") +
  geom_line(aes(y = neurology, x = year, 
                colour = "neurosciences & neurology"), stat="identity") +
  geom_line(aes(y = psychology, x = year, 
                colour = "psychology"), stat="identity") +
  geom_line(aes(y = sociology, x = year, 
                colour = "sociology"), stat="identity") +
  geom_line(aes(y = oncology, x = year, 
                colour = "oncology"), stat="identity") +
  geom_line(aes(y = business, x = year, 
                colour = "business & economics"), stat="identity") +
  labs(title = "Figure 7B. Growth in Diversity-Related Terminology by Select Subject (1990-2017)") + 
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()  
```

When we break the trends in categories down as percentages, the patterns are even more marked. This graph shows that diversity-related terminology has always been highest in genetics & heredity and biochemistry & molecular biology. This terminology has generally increased as a percentage in most other categories, including microbiology, infectious diseases, pharmacology & pharmacy research, psychology, sociology and the behavioral sciences. Some disciplines have shown little to no change such as the neurosciences, business & economics, and oncology. On the other hand, these terms have actually declined markedly in the field of immunology. 

Instead of just selectively visualizing disciplines, we also wanted to know what would happen if we just collapsed the categories into three large categorical buckets: (1) genetics & heredity (i.e. evolution, plant and animal sciences), (2) biomedical studies, and (3) the social and behavioral sciences. As a general note, when one of the original categories did not fit into these three recoded categories (e.g. mathematics or computer science), they were just recoded as "other" and ignored in our final analysis. For a full list of how the original 134 categories were recoded, [you can visit this link](https://docs.google.com/spreadsheets/d/1Gs-aSgXlnwo8rP3wnpMB0kdrU_VS76aDT3DCiI8OmVU/edit#gid=849010328). 

```{r breaking down annual growth of diversity by recoded subject, message = FALSE, warning = FALSE, fig.width=9.5, echo=FALSE}
# loading the recoded subject data 
genetics_heredity <- read_csv("subject_categories.csv") %>% 
  filter(collapsed_subject == "genetics_heredity")
biomedical_studies <- read_csv("subject_categories.csv") %>% 
  filter(collapsed_subject == "biomedical_studies")
social_behavioral <- read_csv("subject_categories.csv") %>% 
  filter(collapsed_subject == "social_behavioral")
other_studies <- read_csv("subject_categories.csv") %>% 
  filter(collapsed_subject == "other")
# recoding the data 
genetics_heredity <- paste(c("\\b(?i)(zxz", genetics_heredity$original_subject, "zxz)\\b"),
                           collapse = "|")
biomedical_studies <- paste(c("\\b(?i)(zxz", biomedical_studies$original_subject, "zxz)\\b"),
                            collapse = "|")
social_behavioral <- paste(c("\\b(?i)(zxz", social_behavioral$original_subject, "zxz)\\b"),
                           collapse = "|")
other_studies <- paste(c("\\b(?i)(zxz", other_studies$original_subject, "zxz)\\b"), 
                       collapse = "|")
```

```{r figure 8a, fig.width=9.5, warning=FALSE, echo=FALSE}
# recoding the subject categories and graphing changes over time 
pop_terms_bysubject <- pop_terms_bysubject %>% 
  mutate(category = ifelse(test = str_detect(string = subject, 
                           pattern = genetics_heredity), 
                           yes = "genetics & heredity", no = "")) %>% 
  mutate(category = ifelse(test = str_detect(string = subject, 
                           pattern = biomedical_studies), 
                           yes = "biomedical studies", no = category)) %>% 
  mutate(category = ifelse(test = str_detect(string = subject, 
                           pattern = social_behavioral), 
                           yes = "social & behavioral sciences", no = category)) %>%
  mutate(category = ifelse(test = str_detect(string = subject, 
                           pattern = other_studies), 
                           yes = "other subjects", no = category)) 
# graphing trends 
pop_terms_sbjrc <- pop_terms_bysubject %>%
  group_by(year) %>% count(term, category, sort = TRUE) %>% ungroup() %>% 
  filter(term == "all population terms") %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = category), mstat="identity") +
  labs(title = "Figure 8A. Growth in Diversity-Related Terminology by Collapsed Category (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

While the growth of diversity-related terminology rises in the social & behavioral sciences and genetics & heredity research, these trends are clearly much not as intense as the increase we see in the biomedical sciences. But maybe this is just a byproduct of the majority of our categories being biomedical? Let's check the percentages...

```{r figure 8b, fig.width=9.5, warning=FALSE, echo=FALSE}
# total articles each year 
bysbjrc_prc_counts <- text_data %>% group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n) 
# articles with term mentioned each year 
bysbjrc_prc_counts <- pop_terms_bysubject %>% filter(diversity == 1) %>% 
  group_by(category, year) %>% summarise(cnt_diversity = n_distinct(id)) %>% 
  right_join(bysbjrc_prc_counts, by = "year") %>% 
  mutate(prc_diversity = round(cnt_diversity / total * 100, digits = 2))
# wrangling all cats into a df 
bysbjrc_graph <- bysbjrc_prc_counts %>% filter(category == "genetics & heredity") %>% 
  mutate(genetics_heredity = prc_diversity) 
bysbjrc_graph <- bysbjrc_prc_counts %>% filter(category == "biomedical studies") %>% 
  mutate(biomedical_studies = prc_diversity) %>% right_join(bysbjrc_graph, by = "year")
bysbjrc_graph <- bysbjrc_prc_counts %>% filter(category == "social & behavioral sciences") %>% 
  mutate(social_behavioral = prc_diversity) %>% right_join(bysbjrc_graph, by = "year")
bysbjrc_graph <- bysbjrc_prc_counts %>% filter(category == "other subjects") %>% 
  mutate(other_studies = prc_diversity) %>% right_join(bysbjrc_graph, by = "year")
# graphing percentages by collapsed cats 
bysbjrc_graph <- bysbjrc_graph %>% ggplot() + 
  geom_line(aes(y = genetics_heredity, x = year, colour = "genetics & heredity"), stat="identity") +
  geom_line(aes(y = biomedical_studies, x = year, colour = "biomedical studies"), stat="identity") +
  geom_line(aes(y = social_behavioral, x = year, colour = "social & behavioral sciences"), stat="identity") +
  geom_line(aes(y = other_studies, x = year, colour = "other subjects"), stat="identity") +
  labs(title = "Figure 8B. Growth in Diversity-Related Terminology by Collapsed Category (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

Here, we see that the percentages generally reflect the same patterns as the raw counts. In other words, diversity-related terminology has always been highest in the biomedical sciences, but has definitely increased as a percentage over the course of the 2000s. We see similar growth trends in genetics & heredity research as well as social & behavioral research. While social & behavioral research has continued to increase up to 2017, it seems as if the other collapsed categories are tapering off (or even declining) over the past five years.  

### Conclusion 

Overall, this document shows a rise in the use of "diversity" across scientific research. We see a 10-fold increase across the 1990s and 2000's, which mostly occurs in research deriving from Westernized biomedical scientific contexts. Our future analyses will examine more what implications this has for the use of diversity in and outside of that domain. 

### References 

Aria, M., & Cuccurullo, C. (2017). bibliometrix: An R-tool for comprehensive science mapping analysis. *Journal of Informetrics*, 11(4), 959-975.

Epstein, S. (2008). *Inclusion: The Politics of Difference in Medical Research.* University of Chicago Press.

Kramer, B. (2019). *Molecularization at the Intersections: Testosterone, Prostate Cancer and the Construction of Racial Difference.* Doctoral dissertation, Rutgers University-School of Graduate Studies.

Lee, C. (2009). Race and ethnicity in biomedical research: How do scientists construct and explain differences in health?. *Social Science & Medicine*, 68(6), 1183-1190.

M'charek, A. (2013). Beyond fact or fiction: On the materiality of race in practice. *Cultural Anthropology*, 28(3), 420-442.

Oh, S. S., Galanter, J., Thakur, N., Pino-Yanes, M., Barcelo, N. E., White, M. J., et al. (2015). Diversity in clinical and biomedical research: A promise yet to be fulfilled. *PLoS Medicine*, 12(12), e1001918.

Panofsky, A., & Bliss, C. (2017). Ambiguity and scientific authority: Population classification in genomic science. *American Sociological Review*, 82(1), 59-87.

Silge, J., & Robinson, D. (2016). tidytext: Text Mining and Analysis Using Tidy Data Principles in R. *Journal of Open Source Software*, 1(3), 37.


