---
title: "The Growth of Diversity"
author: "Brandon L. Kramer (UVA, Biocomplexity Institute) & Catherine Lee (Rutgers, Sociology)"
output: html_document 
---

### Project Overview

Over the course of the 25 years, there have been persistent calls from both governmental and scientific circles to include more diverse populations in biomedical research (Epstein 2008). Among the many initatives that have been put in place since, the National Institute of Health (NIH) spearheaded this process mandating that all federally funded clinical research work to include women and racial/ethinic minorities in their research, citing concerns over population-specific diseases and potential side-effects that may only impact certain groups (Oh et al. 2015). In this project, we are interested in researching how this policy and other inclusion-based initiatives have contributed to the growth of diversity-related terminiology in biomedical research. To do this, we will be using aspects of computational text analysis to outline the emergence of diversity-related terms and how they vary by some common co-variates such as geography and discipline. 

### Data Overview

In our analyses below, we are examining the growth of in the use of the term "diversity". To do this, we drew from the MEDLINE database in Web of Science, using the search terms ["TS=(diversity)" from 1990-2017 for human research only](https://docs.google.com/spreadsheets/d/1RNpPAyowuDfA5p8Sx7nEIw1a7Z8VWFoMFc3WPQ0QELw/edit#gid=933487936). This search provided 71,528 total results, which we extracted using the bibliometrix package in R (Aria and Cuccurillo 2017). Next, we converted the abstracts of these articles to a text corpus and then used tidytext - a package designed for computational text analysis in the open-source software R (Silge and Robinson 2016) - to analyze patterns with the abstracts of these data. Below is a summary of the results, but the replication code can be found on [Brandon's Github page](github.com/brandonleekramerdiversity).

```{r setup, include = FALSE}
rm(list = ls())
#knitr::opts_knit$set(root.dir = "~/Documents/Diversity/Data")
knitr::opts_knit$set(root.dir = "C:/Users/bkram/CloudStation/Diversity/Data")
```

### Overall Growth of Publications 

In this first chunk, we load our data and examine the overall growth of articles from our search query. As we can see, there is a pretty sizable growth in scientific literature that uses the term diversity - from about 500 times in 1990 to over 5000 in 2017.  

```{r load packages, message = FALSE, results = FALSE, echo=FALSE}
# loading packages 
for (pkg in c("tidyverse", "igraph", "tidytext", "ggplot2", "ggraph", "widyr", "plotly", "stringi", "networkD3")) {library(pkg, character.only = TRUE)}
# loading the .csv file 

text_data <- read_csv("biomedical_corpus.csv") %>% 
  rowid_to_column(var = "id") %>% 
  rename(authors = AU, title = TI, publication = SO, #author_keywords = DE,
         abstract = AB, references = CR, year = PY, times_cited = TC) %>% 
  select(id, authors, title, year, publication, abstract, references, year, times_cited) %>% 
# filtering 2018 articles because they seem to be incomplete 
  filter(year != "2018") 
  
```
```{r wos study growth, fig.width=9.5, echo=FALSE}
# checking to see how the overall data looks 
text_data %>% 
  group_by(year) %>% 
  count(year, sort = TRUE) %>% 
  arrange(year)

pubs_by_year <- text_data %>% 
  group_by(year) %>% count(year, sort = TRUE) %>% ungroup() %>% 
  ggplot() + geom_line(aes(y = n, x = year), stat="identity") + 
  labs(title = "Figure 1. Growth in Diversity-Related Publications from 1990-2017",
       caption = "Data Source: Web of Science") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank()); ggplotly()
```

### Word Frequencies Over Time

Next, we want to look at word frequencies by year in the literature. Note that we did remove some frequently occurring words that are not really relevant to our dataset (i.e. custom stopwords), but these do not systematically alter our results. 

```{r text analysis of abstracts, messages = FALSE, warning=FALSE, echo=FALSE}
# tokenizing the abstract data into words 
abstract_data <- text_data %>% 
  unnest_tokens(word, abstract) %>% 
  anti_join(stop_words) %>% 
  drop_na(word)
# adding custom set of stopwords 
my_stopwords <- tibble(word = c(as.character(1:14), 
                                "1", "2", "3", "4", "5", "6", "7", "8", "9", "10", 
                                "rights", "reserved", "copyright", "elsevier"))
abstract_data <- abstract_data %>% anti_join(my_stopwords); rm(my_stopwords)
# most frequent word count in abstracts 
abstract_data %>%
  count(word, sort = TRUE)
```

```{r abstract word count, messages = FALSE, warning=FALSE, include=FALSE}
# looking at word frequencies by year 
abstract_data %>%
  group_by(year) %>% 
  count(word, sort = TRUE) %>% ungroup()
```

### Growth in General Population Terms Over Time

Now that we have the basics out of the way, we want to look at how the most relevant words vary over time. We chose to include words like diversity and its various sub-types (i.e. genetic, cultural, and population) as well as racially-specific and ethnically-specific terms. 

```{r figure 2a, fig.width=9.5, echo=FALSE}
# recoding race*, ethnicity* and cultural* terminology
general_pop_terms <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(race|racial|racially)\\b"), 
                       yes = "racial*", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(ethnic|ethnicity|ethnically)\\b"), 
                       yes = "ethnic*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(culture|cultural|culturally)\\b"), 
                       yes = "cultural*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern =
  "\\b(?i)(ancestry|ancestries|ancestral|ancestor|ancestors|descent|descendent|descendents)\\b"), 
                       yes = "ancestry*", no = term)) %>% 
# grouping by year and filter relevant population terms 
  group_by(year) %>% count(term, sort = TRUE) %>% ungroup() %>% 
  filter(term == "diversity" | term == "cultural*" | term == "genetic" | term == "population*" | 
           term == "racial*" | term == "ethnic*" | term == "ancestry*") %>% 
# graphing terms over time 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), stat="identity") + 
  labs(title = "Figure 2A. Growth in General Diversity-Related Terms (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

As the graph shows, the rise of diversity does not necessarily mean that researchers focus on race or ethnicity. This could mean that diversity is being used as a catch-all in the scientific literature (i.e. that the multiplicity of the term makes it mean anything and everything) or that diversity is most used in fields like immunology or oncology. We will explore that hypothesis a bit more below. Let's take a look at our first graph.

```{r figure 2b, fig.width=9.5, echo=FALSE}
# recoding race*, ethnicity* and cultural* terminology
general_pop_terms <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(race|racial|racially)\\b"), 
                       yes = "racial*", no = word)) %>% 
  mutate(di_racial = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(racial*)\\b"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(ethnic|ethnicity|ethnically)\\b"), 
                       yes = "ethnic*", no = term)) %>% 
  mutate(di_ethnic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(ethnic*)\\b"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(culture|cultural|culturally)\\b"), 
                       yes = "cultural*", no = term)) %>% 
  mutate(di_cultural = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(cultural*)\\b"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population*", no = term)) %>% 
  mutate(di_population = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(population*)\\b"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern =
  "\\b(?i)(ancestry|ancestries|ancestral|ancestor|ancestors|descent|descendent|descendents)\\b"), 
                       yes = "ancestry*", no = term)) %>% 
  mutate(di_ancestry = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(ancestry*)\\b"), yes = 1, no = 0)) %>% 
  mutate(di_diversity = ifelse(test = str_detect(string = term, 
                       pattern = "diversity"), yes = 1, no = 0)) %>% 
  mutate(di_genetic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(genetic)\\b"), yes = 1, no = 0))

# total articles each year 
gen_pop_prc_counts <- text_data %>% 
  group_by(year) %>% 
  count(year) %>% 
  ungroup() %>% 
  rename(total = n) 
# articles with term mentioned each year 
gen_pop_prc_counts <- general_pop_terms %>% 
  filter(di_racial == 1) %>% group_by(year) %>% 
  summarise(cnt_racial = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>%  
  mutate(prc_racial = round(cnt_racial / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_ethnic == 1) %>% 
  group_by(year) %>% summarise(cnt_ethnic = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_ethnic = round(cnt_ethnic / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_cultural == 1) %>% 
  group_by(year) %>% summarise(cnt_cultural = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_cultural = round(cnt_cultural / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_population == 1) %>% 
  group_by(year) %>% summarise(cnt_population = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_population = round(cnt_population / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_ancestry == 1) %>% 
  group_by(year) %>% summarise(cnt_ancestry = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_ancestry = round(cnt_ancestry / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_diversity == 1) %>% 
  group_by(year) %>% summarise(cnt_diversity = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_diversity = round(cnt_diversity / total * 100, digits = 2))
gen_pop_prc_counts <- general_pop_terms %>% filter(di_genetic == 1) %>% 
  group_by(year) %>% summarise(cnt_genetic = n_distinct(id)) %>% 
  right_join(gen_pop_prc_counts, by = "year") %>% 
  mutate(prc_genetic = round(cnt_genetic / total * 100, digits = 2))

# graphing terms over time 
gen_pop_prc_counts <- gen_pop_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_ancestry, x = year, colour = "ancestry*"), stat="identity") +
  geom_line(aes(y = prc_cultural, x = year, colour = "cultural*"), stat="identity") +
  geom_line(aes(y = prc_diversity, x = year, colour = "diversity"), stat="identity") +
  geom_line(aes(y = prc_ethnic, x = year, colour = "ethnic*"), stat="identity") +
  geom_line(aes(y = prc_genetic, x = year, colour = "genetic"), stat="identity") +
  geom_line(aes(y = prc_population, x = year, colour = "population*"), stat="identity") +
  geom_line(aes(y = prc_racial, x = year, colour = "racial*"), stat="identity") +
  labs(title = "Figure 2B. Proportional Growth in General Diversity-Related Terms (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly() 
```

While Figure 2A looks at the raw counts of words mentioned across our dataset, we also want to know what percentage of abstracts that these terms are mentioned in. Figure 2B shows how often general diversity-related terms occur within the corpus of abstracts. As you can see, "diversity" still occurs most often, which is not surprising given our search terms bias the inclusion of that term in a big way. (Note that the search terms include any entry that has diversity in its title, abstract or keywords, which is why we don't see diversity included in 100% of our dataset in any given year.) 

What is interesting about that term is that "diversity", as a term that arises in a certain percentage of scientfic abstracts, actually falls through the early 1990s before rebounding and consistently rising thereafter. Note that the NIH's Health Revitalization Act was instituted in 1993, which may have contributed to this rise of diversity-related terminology. Population, genetic and cultural follow similar upward trends over time, though the rise is much more subdued. On the other hand, race, ancestry and ethnicity all increase as a percentage until the mid-2000's before declining slightly thereafter. Our analyses do not address why these declines occur, though some of the results below suggest that the use of continental and national terms overtake the use of race and ethnicity after 2003, which also happens to be around the time the Human Genome Project was completed. 

### Growth in US-Specific Population Terms Over Time

Now we want to look at all [US-specific population terminology](https://www.census.gov/topics/population/race/about.html). This is not much different than what we have already done except for the small technical problem of tidytext separating hyphenated words in our original analysis (e.g. African-Americans). While the package quanteda has an option to unnest tokens while ignoring hyphenated words, quanteda takes a lot of RAM and ushers in other potential methodological inconsistencies compared to the approach used above. Instead, we are going to create bigrams of key US-specific terms in [tidytext](https://www.tidytextmining.com/ngrams.html) and then add them back into our larger dataset for trend anaylsis. 

```{r figure 3a, fig.width=9.5, echo=FALSE}
# first we need to recode variations of us-specific words (an alternative to stemming) 
us_pop_counts <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word,
                       pattern = "\\b(?i)(caucasian|caucasians)\\b"), yes = "caucasian", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(americans)\\b"), yes = "american", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(hispanics)\\b"), yes = "hispanic", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(indians)\\b"), yes = "indian", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(islanders)\\b"), yes = "islander", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(alaska|alaskans)\\b"), yes = "alaskan", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(hawaii|hawaiians)\\b"), yes = "hawaiian", no = term)) %>%
  select(-word) %>% rename(word = term) %>% 
# create a bigram of us-specific words 
  unnest_tokens(bigram, word, token = "ngrams", n = 2) %>% 
  separate(bigram, c("word1", "word2"), sep = " ") %>%          
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>% 
# unite the birgrams together with a hyphen
  unite(word, word1, word2, sep = "-") %>% 
  filter(word == "african-american" | word == "mexican-american" |
         word == "american-indian" | word == "native-american" | 
         word == "alaskan-native" | word == "native-hawaiian" | 
         word == "pacific-islander" | word == "racial*" | word == "ethnic*" ) %>% 
  count(word, year, sort = TRUE) 
# combining the hyphenated words to our full dataset 
us_pop_counts <- abstract_data %>% 
  group_by(year) %>% count(word, sort = TRUE) %>% ungroup() %>% 
  select(word, year, n) %>% bind_rows(us_pop_counts) 
# graphing us-specific terms over time 
us_pop_terms <- us_pop_counts %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word,
                       pattern = "\\b(?i)(black|blacks|african-american)\\b"), 
                       yes = "black*", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(white|caucasian-american)\\b"), 
                       yes = "white*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(hispanic|mexican-american)\\b"), 
                       yes = "hispanic*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(native-american|american-indian|alaskan-native)\\b"), 
                       yes = "native-american*", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(pacific-islander|native-hawaiian)\\b"), 
                       yes = "pacific-islander*", no = term)) %>%   
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(race|racial|racially)\\b"), 
                       yes = "racial*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(ethnic|ethnicity|ethnically)\\b"), 
                       yes = "ethnic*", no = term)) %>% 
  filter(term == "black*" | term == "white*" | term == "hispanic*" | 
         term == "native-american*" | term == "pacific-islander*" | 
         term == "diversity" | term == "racial*" | term == "ethnic*") %>%  
  group_by(term, year) %>% summarise(n = sum(n)) %>% arrange(-n) %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), stat="identity") + 
  labs(title = "Figure 3A. Growth in US-Specific Diversity-Related Terms (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

To make this graph of all US-specific terminology, we recoded all hyphenated words that we consider US-specific population terms, including African-American, Mexican-American, American-Indian, Native American, Alaskan Indian, and Pacific Islander. We recoded these variations into the five main racial/ethnic categories used on the US Census (White, Black, Hispanic, Native American, and Pacific Islander) and included them alongside all the variations of race and ethnicity. We also included our baseline term diversity to help compare the growth of US-specific terms to our other analyses. As we can see here, there is not a ton of growth in US-specific diversity-related terminology over time (at least in comparison with the term diversity). For those that the growth of diversity is largely related to an increased proclivity for scientists to study race/ethnicity in accordance with a US-specific research program, this graph essentially shows that this hypothesis is pretty weakly supported. 

```{r figure 3b, fig.width=9.5, echo=FALSE}
# first we need to recode variations of us-specific words (an alternative to stemming) 
us_pop_counts <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word,
                       pattern = "\\b(?i)(caucasian|caucasians)\\b"), yes = "caucasian", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(americans)\\b"), yes = "american", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(hispanics)\\b"), yes = "hispanic", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(indians)\\b"), yes = "indian", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(islanders)\\b"), yes = "islander", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(alaska|alaskans)\\b"), yes = "alaskan", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(hawaii|hawaiians)\\b"), yes = "hawaiian", no = term)) %>%
  select(-word) %>% rename(word = term) %>% 
# create a bigram of us-specific words 
  unnest_tokens(bigram, word, token = "ngrams", n = 2) %>% 
  separate(bigram, c("word1", "word2"), sep = " ") %>%          
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>% 
# unite the birgrams together with a hyphen
  unite(word, word1, word2, sep = "-") %>% 
  filter(word == "african-american" | word == "mexican-american" |
         word == "american-indian" | word == "native-american" | 
         word == "alaskan-native" | word == "native-hawaiian" | 
         word == "pacific-islander" | word == "racial*" | word == "ethnic*" ) %>% 
  count(word, id, year, sort = TRUE) 
# combining the hyphenated words to our full dataset 
us_pop_counts <- abstract_data %>% 
  group_by(year) %>% count(word, id, sort = TRUE) %>% ungroup() %>% 
  select(word, id, year, n) %>% bind_rows(us_pop_counts) 
# graphing us-specific terms over time 
us_pop_counts <- us_pop_counts %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word,
                       pattern = "\\b(?i)(black|blacks|african-american)\\b"), 
                       yes = "black*", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(white|caucasian-american)\\b"), 
                       yes = "white*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(hispanic|mexican-american)\\b"), 
                       yes = "hispanic*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(native-american|american-indian|alaskan-native)\\b"), 
                       yes = "native-american*", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(pacific-islander|native-hawaiian)\\b"), 
                       yes = "pacific-islander*", no = term)) %>%   
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(race|racial|racially)\\b"), 
                       yes = "racial*", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = us_pop_counts$word, 
                       pattern = "\\b(?i)(ethnic|ethnicity|ethnically)\\b"), 
                       yes = "ethnic*", no = term)) %>% 
  mutate(black = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(black*)\\b"), yes = 1, no = 0)) %>% 
  mutate(white = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(white*)\\b"), yes = 1, no = 0)) %>% 
  mutate(hispanic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(hispanic*)\\b"), yes = 1, no = 0)) %>% 
  mutate(native_american = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(native-american*)\\b"), yes = 1, no = 0)) %>% 
  mutate(pacific_islander = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(pacific-islander*)\\b"), yes = 1, no = 0)) %>% 
  mutate(racial = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(racial*)\\b"), yes = 1, no = 0)) %>% 
  mutate(ethnic = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(ethnic*)\\b"), yes = 1, no = 0))  

# total articles each year 
us_pop_prc_counts <- text_data %>% group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n) 
# articles with term mentioned each year 
us_pop_prc_counts <- us_pop_counts %>% filter(black == 1) %>% 
  group_by(year) %>% summarise(cnt_black = n_distinct(id)) %>% 
  left_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_black = round(cnt_black / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% filter(white == 1) %>% 
  group_by(year) %>% summarise(cnt_white = n_distinct(id)) %>% 
  left_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_white = round(cnt_white / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% filter(hispanic == 1) %>% 
  group_by(year) %>% summarise(cnt_hispanic = n_distinct(id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_hispanic = round(cnt_hispanic / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% filter(native_american == 1) %>% 
  group_by(year) %>% summarise(cnt_native_american = n_distinct(id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_native_american = round(cnt_native_american / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% filter(pacific_islander == 1) %>% 
  group_by(year) %>% summarise(cnt_pacific_islander = n_distinct(id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_pacific_islander = round(cnt_pacific_islander / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% filter(racial == 1) %>% 
  group_by(year) %>% summarise(cnt_racial = n_distinct(id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_racial = round(cnt_racial / total * 100, digits = 2))
us_pop_prc_counts <- us_pop_counts %>% filter(ethnic == 1) %>% 
  group_by(year) %>% summarise(cnt_ethnic = n_distinct(id)) %>% 
  right_join(us_pop_prc_counts, by = "year") %>% 
  mutate(prc_ethnic = round(cnt_ethnic / total * 100, digits = 2))

# graphing terms over time 
us_pop_prc_counts <- us_pop_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_black, x = year, colour = "black*"), stat="identity") +
  geom_line(aes(y = prc_white, x = year, colour = "white*"), stat="identity") +
  geom_line(aes(y = prc_hispanic, x = year, colour = "hispanic*"), stat="identity") +
  geom_line(aes(y = prc_native_american, x = year, colour = "native-american*"), stat="identity") +
  geom_line(aes(y = prc_pacific_islander, x = year, colour = "pacific-islander*"), stat="identity") +
  geom_line(aes(y = prc_ethnic, x = year, colour = "ethnic*"), stat="identity") +
  geom_line(aes(y = prc_racial, x = year, colour = "racial*"), stat="identity") +
  labs(title = "Figure 3B. Growth in US-Specific Diversity Terms, By Percentage (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly() 
```

When we look at those trends as terms that occur with in a percentage of articles each year, the hypothesis that race/ethnicity is growing in use is even less viable. We can see that US-specific population terminology increases throughout the 1990s and early-2000s before declining after that. As I mentioned above, it is unclear what exactly might be causing this, but one possibility is that national and continental terms replace US-specific research as international collaboration increased after the completion of the Human Genome Project. 

### The Multiplicity of Population Terminology

While that graph was quite informative, we do not see much variation in how race/ethinicity vary over time in comparison with terms like diversity. One possibility is that variation in population terms is "diluted" by both the multiplicity of population terminology and the ambiguity of how those terms are used (M'Charek 2013; Panofsky and Bliss 2017). Thus, the next logical step is to collapse all possible population terms into one category and graph that over time. Both Lee (2009) and Kramer (2019) as well as others (e.g. Panofsky and Bliss 2017) have found that biomedical researchers continue to use population-specific terminology to reinforce the notion of population differences in various biological markers. Below, we have collapsed several population-specific terms into one category based. [This list of terms](https://docs.google.com/spreadsheets/d/1Gs-aSgXlnwo8rP3wnpMB0kdrU_VS76aDT3DCiI8OmVU/edit#gid=0) which were developed out of Kramer's (2019) dissertation work. While we won't claim that this is an exhaustive list of all the population terms that exist in the world, it is a fairly comprehensive list of over 2,100 terms (also see below in text networks section). Let's take a look how much the use of all these population-specific terms grew over time.

```{r collapsing population specific terms, message = FALSE, results = FALSE, warning = FALSE, echo=FALSE}
# pulling our list of population
all_pop_terms <- read_csv("population_terms.csv") 
all_pop_terms <- paste(c("\\b(?i)(zcx", all_pop_terms$term, "zxc)\\b"), collapse = "|")
```

```{r figure 4a, fig.width=9.5, echo=FALSE}
all_pop_terms_graph <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population*", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = all_pop_terms), 
                       yes = "all population terms", no = term)) %>% 
  group_by(year) %>% count(term, sort = TRUE) %>% ungroup() %>% 
  filter(term == "diversity" | term == "genetic" | 
         term == "population*" | term == "all population terms") %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), stat="identity") + 
  labs(title = "Figure 4A. Growth in Collapsed Diversity-Related Terminology (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

Interesting! After we collapsed all of the population terms together into one category, we see a dramatic growth that goes well beyond even the term "diversity" that was so high to begin with. This give some support to the notion that multiplicity and ambiguity could be mechanisms that drive the growth of diversity-related terms. More specifically, following M'Charek (2013), we are arguing that race is a multiplicity; it is a discourse that can link to *various* material referents depending on the context it is embedded within. The definition of race can be dependent on the technologies or researchers that are enacting the concept to carry out specific ends. Because of its multiplicitous nature, race also becomes ambiguous and uncertain. Race can be many things, which obscures the precision of any specific claims that can be made about it. Panofsky and Bliss (2017) find that population terms grow in use over time in the context of biomedical studies, but there use is often shifting and imprecise. The authors argue that this ambiguity seems to even reinforce authority in some scientific communities. Building on that work, we do see evidence that variability in population terms at multiple levels does increase over time. We see the same trends when we look the growth of all population claims as a percentage as well (below).

```{r figure 4b, fig.width=9.5, echo=FALSE}
all_pop_terms_trends <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population*", no = word)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = all_pop_terms), 
                       yes = "all population terms", no = term)) %>% 
  mutate(population = ifelse(test = str_detect(string = term, 
                       pattern = "\\b((?<!all )population*)\\b"), yes = 1, no = 0)) %>% 
  mutate(all_pop_terms = ifelse(test = str_detect(string = term, 
                       pattern = "\\b(all population terms)\\b"), yes = 1, no = 0)) 
# total articles each year 
all_pop_prc_counts <- text_data %>% group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n) 
# articles with term mentioned each year 
all_pop_prc_counts <- all_pop_terms_trends %>% filter(population == 1) %>% 
  group_by(year) %>% summarise(cnt_population = n_distinct(id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_population = round(cnt_population / total * 100, digits = 2))
all_pop_prc_counts <- all_pop_terms_trends %>% filter(all_pop_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_all_pop = n_distinct(id)) %>% 
  right_join(all_pop_prc_counts, by = "year") %>% 
  mutate(prc_all_pop = round(cnt_all_pop / total * 100, digits = 2))
# graphing the patterns over time
all_pop_prc_counts <- all_pop_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_population, x = year, colour = "population*"), stat="identity") +
  geom_line(aes(y = prc_all_pop, x = year, colour = "all population terms"), stat="identity") + 
  labs(title = "Figure 4B. Growth in Collapsed Diversity Terminology, By Percentage (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

Now, let's see what the growth of all these population terms when we break them into general, national, continental, and ethnic-specific variations over time. Here, general refers to terms like race, ethnicity, and other terms related broadly to racial terms. National and continental terms are self-explanatory. Ethnic-specific terms refers to all possible ethnic terms we could find in the course of generating these terms. That said, we still have some work to do on refining these various terms before publication.

```{r breaking apart collapsed population terms, message = FALSE, warning = FALSE, echo=FALSE}
# pulling in concatenated population terms from our dataset 
general_pop_terms <- read_csv("population_terms.csv") %>% 
  filter(sub_category == "population_general")
us_specific_terms <- read_csv("population_terms.csv") %>% 
  filter(sub_category == "us_specific")
continental_terms <- read_csv("population_terms.csv") %>% 
  filter(category == "continental" | category == "subcontinental")
ling_religious_terms <- read_csv("population_terms.csv") %>% 
  filter(category == "linguistic_religious")
national_terms <- read_csv("population_terms.csv") %>% 
  filter(category == "national")
south_american_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "south_america")
african_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "africa")
north_american_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "north_america")
european_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "europe")
asian_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "asia")
all_ethnic_groups <- read_csv("population_terms.csv") %>% 
  filter(category == "south_america" | category == "africa" | category == "asia" | 
         category == "north_america" | category == "europe")
# creating detection string patterns for all the categories
general_pop_terms <- paste(c("\\b(?i)(zxz", 
                     general_pop_terms$term, "zxz)\\b"), collapse = "|")
us_specific_terms <- paste(c("\\b(?i)(zxz", 
                     us_specific_terms$term, "zxz)\\b"), collapse = "|")
continental_terms <- paste(c("\\b(?i)(zxz", 
                     continental_terms$term, "zxz)\\b"), collapse = "|")
ling_religious_terms <- paste(c("\\b(?i)(zxz", 
                     ling_religious_terms$term, "zxz)\\b"), collapse = "|")
national_terms <- paste(c("\\b(?i)(zxz", 
                     national_terms$term, "zxz)\\b"), collapse = "|")
south_american_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     south_american_ethnic_groups$term, "zxz)\\b"), collapse = "|")
african_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     african_ethnic_groups$term, "zxz)\\b"), collapse = "|")
north_american_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     north_american_ethnic_groups$term, "zxz)\\b"), collapse = "|")
european_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     european_ethnic_groups$term, "zxz)\\b"), collapse = "|")
asian_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     asian_ethnic_groups$term, "zxz)\\b"), collapse = "|")
all_ethnic_groups <- paste(c("\\b(?i)(zxz", 
                     all_ethnic_groups$term, "zxz)\\b"), collapse = "|")
```

```{r figure 5a, fig.width=9.5, echo=FALSE}
# graphing the population terms into various categories 
pop_terms_bycats <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = general_pop_terms), 
                       yes = "general population terms", no = word)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = us_specific_terms), 
                       yes = "us-specific terms", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = continental_terms), 
                       yes = "continental terms", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = ling_religious_terms), 
                       yes = "linguistic & religious terms", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = national_terms), 
                       yes = "national terms", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = south_american_ethnic_groups), 
                       yes = "south american ethnic groups", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = african_ethnic_groups), 
                       yes = "african ethnic groups", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = north_american_ethnic_groups), 
                       yes = "north american ethnic groups", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = european_ethnic_groups), 
                       yes = "european ethnic groups", no = term)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = asian_ethnic_groups), 
                       yes = "asian ethnic groups", no = term)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population*", no = term)) %>% 
  group_by(year) %>% count(term, sort = TRUE) %>% ungroup() %>% 
  filter(term == "diversity" | term == "population*" | 
         term == "general population terms" | term == "us-specific terms" |
         term == "continental terms" | term == "linguistic & religious terms" | 
         term == "national terms" | term == "south american ethnic groups" | 
         term == "african ethnic groups" | term == "north american ethnic groups" | 
         term == "european ethnic groups" | term == "asian ethnic groups" ) %>% 
  ggplot() + geom_line(aes(y = n, x = year, colour = term), stat="identity") + 
  labs(title = "Figure 5A. Growth in Diversity-Related Terminology by Category (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

As we can see here, the majority of growth in population-specific terms is because of an increased proclivity to mention nationalities (e.g. Italian, Japanese, Canadian, etc.), continental terms (e.g. Africa, Asia, Europe, or North and South America), and general population terms (e.g. race, ethnicity, Caucasian, African, Asian, etc.). Relative to these other general terms, we do not really see much growth in ethnic groups in any given continent or when we lump all of the different ethnic groups into one category. 

```{r figure 5b, fig.width=9.5, echo=FALSE}
# graphing the population terms into various categories 
pop_terms_bycats <- abstract_data %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = general_pop_terms), 
                       yes = "general population terms", no = word)) %>%
  mutate(general_terms = ifelse(test = str_detect(string = term, 
                       pattern = "general population terms"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = us_specific_terms), 
                       yes = "us-specific terms", no = term)) %>%
  mutate(us_terms = ifelse(test = str_detect(string = term, 
                       pattern = "us-specific terms"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = continental_terms), 
                       yes = "continental terms", no = term)) %>%
  mutate(continental_terms = ifelse(test = str_detect(string = term, 
                       pattern = "continental terms"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = ling_religious_terms), 
                       yes = "linguistic & religious terms", no = term)) %>%
  mutate(ling_rel_terms = ifelse(test = str_detect(string = term, 
                       pattern = "linguistic & religious terms"), yes = 1, no = 0)) %>% 
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = national_terms), 
                       yes = "national terms", no = term)) %>%
  mutate(national_terms = ifelse(test = str_detect(string = term, 
                       pattern = "national terms"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = south_american_ethnic_groups), 
                       yes = "south american ethnic groups", no = term)) %>%
  mutate(south_american_terms = ifelse(test = str_detect(string = term, 
                       pattern = "south american ethnic groups"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = african_ethnic_groups), 
                       yes = "african ethnic groups", no = term)) %>%
  mutate(african_terms = ifelse(test = str_detect(string = term, 
                       pattern = "african ethnic groups"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = north_american_ethnic_groups), 
                       yes = "north american ethnic groups", no = term)) %>%
  mutate(north_american_terms = ifelse(test = str_detect(string = term, 
                       pattern = "north american ethnic groups"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = european_ethnic_groups), 
                       yes = "european ethnic groups", no = term)) %>%
  mutate(european_terms = ifelse(test = str_detect(string = term, 
                       pattern = "european ethnic groups"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = asian_ethnic_groups), 
                       yes = "asian ethnic groups", no = term)) %>% 
  mutate(asian_terms = ifelse(test = str_detect(string = term, 
                       pattern = "asian ethnic groups"), yes = 1, no = 0)) %>%
  mutate(term = ifelse(test = str_detect(string = abstract_data$word, 
                       pattern = "\\b(?i)(population|populations)\\b"), 
                       yes = "population*", no = term)) %>% 
  mutate(population = ifelse(test = str_detect(string = term, 
                       pattern = "\\b((?<!general )population*)\\b"), yes = 1, no = 0)) 

# total articles each year 
bycats_prc_counts <- text_data %>% group_by(year) %>% 
  count(year) %>% ungroup() %>% rename(total = n) 
# articles with term mentioned each year 
bycats_prc_counts <- pop_terms_bycats %>% filter(general_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_general_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_general_terms = round(cnt_general_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(us_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_us_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_us_terms = round(cnt_us_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(continental_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_continental_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_continental_terms = round(cnt_continental_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(ling_rel_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_ling_rel_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_ling_rel_terms = round(cnt_ling_rel_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(national_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_national_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_national_terms = round(cnt_national_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(south_american_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_south_american_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_south_american_terms = round(cnt_south_american_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(african_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_african_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_african_terms = round(cnt_african_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(north_american_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_north_american_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_north_american_terms = round(cnt_north_american_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(european_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_european_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_european_terms = round(cnt_european_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(asian_terms == 1) %>% 
  group_by(year) %>% summarise(cnt_asian_terms = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_asian_terms = round(cnt_asian_terms / total * 100, digits = 2))
bycats_prc_counts <- pop_terms_bycats %>% filter(population == 1) %>% 
  group_by(year) %>% summarise(cnt_population = n_distinct(id)) %>% 
  right_join(bycats_prc_counts, by = "year") %>% 
  mutate(prc_population = round(cnt_population / total * 100, digits = 2))

bycats_prc_counts <- bycats_prc_counts %>% ggplot() + 
  geom_line(aes(y = prc_general_terms, x = year, colour = "general population terms"), stat="identity") +
  geom_line(aes(y = prc_us_terms, x = year, colour = "us-specific terms"), stat="identity") +
  geom_line(aes(y = prc_continental_terms, x = year, colour = "continental terms"), stat="identity") +
  geom_line(aes(y = prc_ling_rel_terms, x = year, colour = "linguistic & religious terms"), stat="identity") +
  geom_line(aes(y = prc_national_terms, x = year, colour = "national terms"), stat="identity") +
  geom_line(aes(y = prc_south_american_terms, x = year, colour = "south american terms"), stat="identity") +
  geom_line(aes(y = prc_african_terms, x = year, colour = "african terms"), stat="identity") +
  geom_line(aes(y = prc_north_american_terms, x = year, colour = "north american terms"), stat="identity") +
  geom_line(aes(y = prc_european_terms, x = year, colour = "european terms"), stat="identity") +
  geom_line(aes(y = prc_asian_terms, x = year, colour = "asian terms"), stat="identity") +
  geom_line(aes(y = prc_population, x = year, colour = "population*"), stat="identity") +
  labs(title = "Figure 5B. Growth in Diversity-Related Terminology, By Category/Percentage (1990-2017)") + 
  theme(axis.title.x = element_blank(), axis.title.y = element_blank(), 
        legend.title = element_blank()); ggplotly()
```

When we look at the trends by percentage, the same national, continental and general population terminology is still most likely to grow over time. The main difference is that as a percentage, these terms taper off after the mid-2000's again, which is also the case with specific continental and ethnic terminology as well as US-specific vernacular. The exception to this plateau effect are national terms, which grow strongly all the way up to the present day. 

### Conclusion 

Overall, this document shows a rise in the use of "diversity" across scientific research. We see a 10-fold increase across the 1990s and 2000's, which mostly occurs in research deriving from Westernized biomedical scientific contexts. Our future analyses will examine more what implications this has for the use of diversity in and outside of that domain. 

### References 

Aria, M., & Cuccurullo, C. (2017). bibliometrix: An R-tool for comprehensive science mapping analysis. *Journal of Informetrics*, 11(4), 959-975.

Epstein, S. (2008). *Inclusion: The Politics of Difference in Medical Research.* University of Chicago Press.

Kramer, B. (2019). *Molecularization at the Intersections: Testosterone, Prostate Cancer and the Construction of Racial Difference.* Doctoral dissertation, Rutgers University-School of Graduate Studies.

Lee, C. (2009). Race and ethnicity in biomedical research: How do scientists construct and explain differences in health?. *Social Science & Medicine*, 68(6), 1183-1190.

M'charek, A. (2013). Beyond fact or fiction: On the materiality of race in practice. *Cultural Anthropology*, 28(3), 420-442.

Oh, S. S., Galanter, J., Thakur, N., Pino-Yanes, M., Barcelo, N. E., White, M. J., et al. (2015). Diversity in clinical and biomedical research: A promise yet to be fulfilled. *PLoS Medicine*, 12(12), e1001918.

Panofsky, A., & Bliss, C. (2017). Ambiguity and scientific authority: Population classification in genomic science. *American Sociological Review*, 82(1), 59-87.

Silge, J., & Robinson, D. (2016). tidytext: Text Mining and Analysis Using Tidy Data Principles in R. *Journal of Open Source Software*, 1(3), 37.


